### 2012-10-31 FIGGUScode.tex
adicionado arquivo de apendice

\chapter{Código do FIGGUS (Finite Groups in Granular and Unit Synthesis}
\label{cap:FIGGUScode}