### 2013-09-19 article/article.tex
fazendo o artigo, ainda na intro

% notas:
% campo harmônico => harmonic field?
% anti-relative => counter-relative? (discussão sobre harmonia funcional)
%    associada relative => associated relative? (ver a discussão para mais...)
% medianas cromáticas => cromatic mediant
% tonalidade => key
% homônica => homonymous

\documentclass[
 aip,
 jmp,
 amsmath,amssymb,
 reprint,
]{revtex4-1}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{multirow}
\usepackage{color}
\usepackage{xcolor}
\usepackage{enumerate}
\usepackage{float}

\newcommand{\massa}{{\large \textsc{massa}}}
\newcommand{\mass}{{\large \textsc{mass}}}
\newcommand{\figgus}{{\large \textsc{figgus}}}

\floatstyle{plain}
\newfloat{twocolequfloat}{t}{zzz}
\floatname{twocolequfloat}{Equation}

\begin{document}
\preprint{AIP/123-QED}

\title{Psychophysics of musical elements in the discrete-time representation of sound}

\author{Renato Fabbri}
 \homepage{http://www.ifsc.usp.br/~fabbri}
 \email{renato.fabbri@gmail.com}
  \affiliation{ 
Instituto de F\'isica de S\~ao Carlos, Universidade de S\~ao Paulo (IFSC/USP)
}

\author{Vilson Vieira da Silva Jr.}
  \homepage{http://automata.cc/}
  \email{vilson@void.cc}
 \altaffiliation[Also at ]{IFSC-USP}

\author{D\'ebora Corr\^ea}
  \homepage{}
  \email{debcris.cor@gmail.com}
 \altaffiliation[Instituto de Ci\^encias Matem\'aticas e de Computa\c{c}\~ao ]{ICMC-USP}

\author{Luciano da Fontoura Costa}
  \homepage{http://cyvision.ifsc.usp.br/~luciano/}
  \email{ldfcosta@gmail.com}
 \altaffiliation[Also at ]{IFSC-USP}

 \author{Osvaldo N. Oliveira Jr.}
  \homepage{www.polimeros.ifsc.usp.br/professors/professor.php?id=4}
  \email{chu@ifsc.usp.br}
 \altaffiliation[Also at ]{IFSC-USP}

\date{\today}
\begin{abstract}

    %The representation of the basic elements of music - such as notes, ornaments and interval structures - in terms of discrete audio signal is often used in software for music creation and design. Nevertheless, there is no unified approach that relates these elements to the sound discrete samples. In this article, each musical element is described by equations in sonic time samples, which are all implemented in scripts within an open source software toolbox, referred to as \massa\ (Music and Audio in Sequences and Samples). The fundamental element, the musical note with duration, volume, pitch and timbre, is related quantitatively to the characteristics of the discrete-time signal. Internal variations, such as tremolos, vibratos and spectral fluctuations, are also considered, which enables the synthesis of notes inspired by real instruments and new sonorities. With this representation of notes, resources are provided for the generation of musical structures, such as rhythmic meter, pitch intervals and cycles. The efficacy of these physical descriptions of basic musical elements was confirmed by the synthesis of small musical pieces within each frame: basic notes, incremented notes and notes in music. It is possible to synthesize whole albums through collage of the scripts and parametrization. 
    %The sample-based analytical description, and the paradigm of open source implementation, enables scientific experiments in precise and trustful ways. In fact, \massa\ has already been employed by external users for diverse purposes. Among these, it is mentioned an acoustic effect recognized by diverse individuals in mailing lists but not found in literature, and uses related to art and education.

Notes, ornaments and intervals are examples of basic elements of music,  their representation as discrete-time digital audio signal plays a central role in software for music creation and design. Nevertheless, there is no systematic relation, in analytical terms, of these musical elements to the sound samples. Such a compendium is important since it enables scientific experiments in precise and trustful ways. In this paper, a comprehensive description of the musical elements within a unified approach is presented. Musical elements, like pitch, duration and timbre are expressed by equations on sample level. This quantitatively relates characteristics of the discrete-time signal to musical qualities. Internal variations, e.g., tremolos, vibratos and spectral fluctuations, are also considered as means to achieve variation within a note. Moreover, the generation of musical structures such as rhythmic meter, pitch intervals and cycles, are reached canonically with notes. The availability of these resources in scripts  (\massa\ - Music and Audio in Sequences and Samples) is also provided in public domain. Authors observe that the implementation of sample-domain analytical results as open source can encourage concise research.
As further illustrated in the paper, \massa\ has already been employed by users for diverse purposes, including acoustics experimentation, art and education. The efficacy of these physical descriptions of basic musical elements was confirmed by the synthesis of small musical pieces.
As shown, it is possible to synthesize whole albums through collage of scripts and parametrization.

\end{abstract}
\pacs{*43.66.-x,43.66.+y,05.10.-a} % PACS, the Physics and Astronomy
\keywords{psychophysics, acoustics, statistics, signal processing, digital audio, music}
\maketitle

\section{\label{sec:level1}Introduction}

Music is commonly defined as the art made by sounds and silences. Sound is the longitudinal wave of mechanical pressure. The frequency bandwidth between $20Hz$ and $20kHz$ is appreciated by the human hearing system with boundaries dependent on the person, climate conditions and sonic characteristics themselves~\cite{Roederer}. Considered the speed of sound as $\approx 343.2 m/s$, these limits corresponds to wavelengths of $\frac{343.2}{20} = 17.16\,m$ and $\frac{343.2}{20000}=17.16\,mm$.

Human perception of sound involves captivation by bones, stomach, ears, transfer functions of head and processing torso and nervous system~\cite{}. Besides that, the ear is a dedicated organ responsible for the capture of these waves. Its mechanism decomposes sound into its sinusoidal spectrum and delivers them to the nervous system. The sinusoidal components are crucial to musical phenomena, as one can perceive the composition of sounds according to a specific musical interest, e.g., tunings and scales.

The representation of sound is commonly called audio (although these terms are often used without distinction). Audio expresses the waveform prominent from the capture of microphones or from synthesis. In order to account for the corresponding digital representation, digital audio is specified by protocols that eases file storage and transferring, usually with low loss of quality. Standard representation of digital audio consists of equally spacing samples by $\lambda_s$ durations in time, where each sample is specified by a sample number of bits. This process is called the Pulse Code Modulation (PCM) representation of sound. A PCM digital sound is characterized by it's sampling frequency $f_s=\frac{1}{\lambda_s}$ (also called the sampling rate), and by a bit depth, which is the number of bits used for representing the amplitude of each sample. With the purposes of illustrating this idea, Figure~\ref{fig:PCM} shows $25$ samples of a PCM audio with $4$ bits each. The $2^4=16$ possible steps for each sample, together with the regular spacing $\lambda_s$ between them, establish what is named as quantization error. The noise prominent of these errors diminishes conforming the spacing also does.

\begin{figure*}
    \centering
        \includegraphics[width=\textwidth]{figures/pcm}
        \caption{Pulse Code Modulation (PCM) audio: an analogical signal is represented by 25 samples with 4 bits each.}
        \label{fig:PCM}
\end{figure*}

The Nyquist theorem~\cite{} says that half of the sampling frequency is the maximum frequency of the signal. Thus, it is necessary to have a sampling frequency at least twice the highest frequency heard by humans, that is, $f_s \geq 2\times 20kHz = 40kHz$. This configures the basis for the use of sampling frequencies like $44.1kHz$ and $48kHz$, standards in Compact Disks (CD) and Broadcast systems (Radio and TV), respectively.

For the average listener -- and a reasonable part of specialists -- the notion of music presupposes rhythmic and pitch organization. I a simple  way, rhythm can be described as recurring pattern produced by notes differing in duration, pause (silence) and intensity. Pitch configures the perceptual attribute that allows sounds to be organized in a frequency-related scale~\cite{}. Music from the twentieth century enlarged this traditional comprehension of music. This occurred in the concert music, specially in the concrete, and in both electronic and electroacoustic styles. On the last decade of the century, it became evident that popular music had also incorporated sounds without defined pitch and temporal organization, to name a few of simple metrics. Even though, the note still stands paradigmatic as a 'fundamental unit' of musical structures and, in practice, it can unfold in sounds that observe this recent developments (**not clear here**).

\subsection{Contributions and paper organization}

This paper purposes to represent musical structures and artifices by their discrete sound characteristics. The results include mathematical relations and their computer program implementations. Despite the general interest and number of knowledge areas involved, there are only a few number of books and computer implementations that tackle the subjects presented in this work. However, they mainly focus on aspects of musical sounds and ways to mimic traditional instruments. A detailed compilation of them is pointed out in the bibliography~\cite{dissertacao}. To the best of the author's knowledge, there is a lack of articles on the topic. Moreover, although current computer implementations use the analytical descriptions presented in this study in a implicitly manner, it seems that there is not a concise and mathematical description of the processes implemented, as they aim here to be further used as libraries for sound and music.

In order to address the concise description of musical elements and structures, the objectives of this paper are:

\begin{enumerate}

\item Present a concise set of relations among musical basic elements and sequences of PCM audio samples.

\item Introduce a framework of sound synthesis with control at sample level and with potential uses in psycho-acoustical experiments and high-fidelity synthesis. The didactic presentation of the content favors its apprehension and usage. It is worthwhile to mention that this subject comprises diverse topics on signal processing, music and psycho-acoustics.

\item Provide the accessibility of the developed framework. The topics presented in this article are implemented as scripts, i.e. small computer programs using accessible technologies for better distribution and validation. This constitute the \massa\ toolbox, available in public domain in an open Git repository~\cite{gitBook}. The scripts are written in Python and make use of external libraries such as Numpy and Scikits/Audiolab, that performs calls to Fortran routines for better computational efficiency. Part of the scripts have been transcribed to JavaScript and native Python with readiness, what favors their use in Web browsers such as Firefox and Chromium~\cite{numpy, audiolab, tutpython, python}. Furthermore, these are all open technologies, that is, published using licenses that allows copy, distribution and use of any part for research and derivatives. This way, the work presented here is available and eases co-authorship processes~\cite{Raymond,Lessig}.

\end{enumerate}

The remaining parts of this work are organized as follows: section~\ref{sec:dicNote} exposes the presence of sinusoid in discrete-time audio and characterizes a basic musical note; the definition and expansion of the musical note as the essential unit of music is further approached in sections~\ref{sec:discNote} and~\ref{sec:internalVar}, respectively; section~\ref{sec:notesMusic} tackles the organization of musical notes in a higher level~\cite{Wisnick,Webern,Lerdhal,Cook,Lacerda}. Musical theory embody diverse topics as psycho-acoustics, cultural manifestations and formalisms. The section~\ref{sec:results} points out these topics as needed and designate external complements~\cite{Zamacois,Schoenberg,microsound}.

The next section is a minimum text in which music elements are presented side-by-side with the discrete-time samples they result. In order to account for validation and sharing, implementations on computer code of these relations and little musical pieces resulting from them are gathered in the \massa\ toolbox, online available.

%Representing musical structures and artifices by it's discrete sound characteristics is the purpose of this work. The results include mathematical relations and it's computer program implementations. Next section exposes the theoretical description, which is implemented as scripts in a one-to-one relation to the equations.

%\subsection{Sound and digital audio}

%Sound is a longitudinal wave of mechanical pressure. The frequency bandwidth between $20Hz$ and $20kHz$ is appreciated by human hearing system with boundaries dependent on the person, climate conditions and sonic characteristics itself~\cite{Roederer}. If considered the speed of sound of $\approx 343.2 m/s$, this limits corresponds to $\frac{343.2}{20} = 17.16\,m$ and $\frac{343.2}{20000}=17.16\,mm$.

%Human perception of sound involves captivation by bones, stomach, ears, transfer functions of head and processing torso and nervous system. Besides that, the ear is a dedicated organ to the capture of this waves. Its mechanism decomposes sound into its sinusoidal spectrum and delivers them to the nervous system. This sinusoidal components are crucial to musical phenomena, as one can observe in the composition of sounds with musical interest and in tunings and scales. Subsection~\ref{subsec:dicNote} exposes the presence of sinusoid in discrete-time audio and characterizes a basic musical note.

%The representation of sound is called audio (although these terms are often used without distinction), and this can be provenient from caption by microphones or from synthesis. Often enough, digital audio is specified by protocols that eases file storage and transferring, in cost of a direct representation or even some loss in quality. Standard representation of digital audio, on the other hand, consists of samples equally spaced by $\lambda_s$ durations in time, with each sample specified by a sample number of bits. This is called the Pulse Code Modulation representation of sound (PCM). A PCM digital sound is characterized by it's sampling frequency $f_s=\frac{1}{\lambda_s}$, also called sampling rate, and bit depth, which is the number of bits used of representing the amplitude of each sample. Figure~\ref{fig:PCM} shows $25$ samples of a PCM audio with $4$ bits each. The $2^4=16$ possible steps for each sample, together with the regular spacing $\lambda_s$ between them, introduces a quantization error. This noise, caused by this errors, diminishes as these spacing diminishes.

%\begin{figure*}
%    \centering
%        \includegraphics[width=\textwidth]{pcm}
%        \caption{Pulse Code Modulation (PCM) audio: an analogical signal is %represented by 25 samples with 4 bits each.}
%        \label{fig:PCM}
%\end{figure*}

%By the Nyquist theorem, it is known that half the sampling frequency is the maximum frequency of the signal. Thus, it is necessary to have a sampling frequency at least twice the highest frequency heard by humans $f_s \geq 2\times 20kHz = 40kHz$. This is the basis for the use of the sampling frequencies $44.1kHz$ and $48kHz$, standards in Compact Disks (CD) and Broadcast systems (Radio and TV), respectively.

%\subsection{Sonic art and musical theory}

%A common definition for music is the art made by sounds and silences. For the average listener -- and a reasonable part of specialists -- the notion of music presupposes rhythmic and pitch organization such as explained in subsection~\ref{subsec:notesMusic}. Music from the twentieth century enlarged this traditional comprehension of music. This occurred in concert music, specially in the concrete, electronic and electroacoustic styles. On the last decade of the century, it was evident that popular music has also incorporated sounds without defined pitch and temporal organization out of simple metrics. Even though, the note stands paradigmatic as a 'fundamental unit' of musical structures and, in practice, it can unfold in sounds that observe this recent developments. The definition and expansion of the musical note as the fundamental unit of music is approached in subsections~\ref{subsec:discNote} and~\ref{subsec:internalVar}, respectively. Subsection~\ref{subsec:notesMusic} tackles the organization of musical notes in a higher level~\cite{Wisnick,Webern,Lerdhal,Cook,Lacerda}.

%Musical theory embody topics as diverse as psycho-acoustics, cultural manifestations and formalisms. The section~\ref{sec:results} point this topics as needed and designate external complements~\cite{Zamacois,Schoenberg,microsound}.

%\subsection{Computational implementation}

%The results presented in this article are implemented as scripts, i.e.\ small computer programs implemented using accessible technologies for better distribution and validation. This constitute the \massa\ toolbox, available in public domain in an open Git repository~\cite{gitBook}. This scripts are written in Python and make use of external libraries Numpy and Scikits/Audiolab that performs calls to Fortran routines for better computational efficiency. Part of this code has been transcribed to JavaScript and native Python with readiness, what points to uses of this contribution in Web browsers such as Firefox and Chromium~\cite{numpy, audiolab, tutpython, python}.

%This are all open technologies, that is, published using licenses that allows copy, distribution and use of any part for research and derivatives. This way, the work here presented is available and eases co-authorship processes~\cite{Raymond,Lessig}.

%\subsection{Objectives}
%\label{subsec:objectives}
%The main goal of this article is to present a concise set of relations among musical basic elements and sequences of PCM audio samples. The next section is a minimum text in which music elements are presented side-by-side with the discrete-time samples they result. As validation and sharing, implementations on computer code of these relations and little musical pieces where gathered the \massa\ toolbox, available online.

%Secondary objectives include presenting a framework of sound synthesis with control at sample level, with potential uses in psychoacoustical experiments and high-fidelity synthesis. The didactic presentation of the content favors use and apprehension on a problem which calls diverse topics to be tackled: signal processing, music and psycho-acoustics, to name just a few.

%\subsection{Related work}
%Due to the general interest, and number of knowledge areas involved, there is a number of books and computer implementations that are of interest or present similarities to what is presented in this work. A more detailed comparison of them is pointed out in the bibliography~\cite{dissertacao}. There is almost no articles which could be found on the topic. In summary, there are computer implementations that use this analytical descriptions implicitly, but there is no such a concise and mathematical description of the processes implemented, as they aim to be libraries for sound and music. There are books on the topic that cover various aspects of effects and physical modeling, but none of them carry a concise description of musical elements and structures, but focus on aspects of musical sounds and ways to mimic traditional instruments.

\section{Characterization of the discrete-time musical note} \label{sec:discNote}

In diverse artistic and theoretical contexts, music is though as being constituted by units called notes, and these units taken as "atoms" that constitute music itself~\cite{Wisnick, Lovelock, Webern}. (****nao entendi****, texto ruim)
Nowadays, these notes are understood as central elements of certain musical paradigms. In a cognitive perspective, the notes are seen as discretized elements that easy and enrich the flow of information through music~\cite{Roederer, Lacerda}.
Canonically, the principal properties of a musical note are duration, volume, pitch and timbre~\cite{Lacerda}. These qualities can be managed quantitatively, dictated by the evenly time spaced sound samples.

All the relations described in this section are implemented at the file \emph{eqs2.1.py} of the \massa\ toolbox. Musical pieces \emph{5 sonic portraits} and \emph{reduced-fi} are also available online in order to corroborate the concepts.

\subsection{Duration}

The sample frequency $f_s$ is defined as the number of samples in each second of the discrete-time signal. Let $T_i=\{t_i\}$ be an ordered set of real samples separated by $\delta_s=1/f_s$ seconds. A musical note of duration $\Delta$ seconds is presented as a sequence of $\lfloor \Delta . f_s \rfloor $ samples. That is, taken the integer part of the multiplication, it is admitted an error of at most $-\lambda_a$ seconds, which are usually fine for musical purposes. Considering $f_s=44.1kHz \;\;\Rightarrow\;\;\lambda_s=\frac{1}{44100}\approx 23$ microseconds. It is reasonable to state that:

\begin{equation}\label{eq:dur}
T_{i}^{\Delta}={\{t_i\}}_{i=0}^{\lfloor \Delta . f_a \rfloor -1}
\end{equation}

Being $\Lambda=\lfloor \Delta . f_a \rfloor$ the number of samples in the sequence, thus $T_i=\{t_i\}_0^{\Lambda-1}$.

\subsection{Volume}\label{subsec:volume}

The sensation of sound volume depends on reverberation and harmonic distribution, among other characteristics described on section~\ref{sec:varInternas}. One can get volume variations through the potency of the wave~\cite{Chowning}:

\begin{equation}\label{eq:potencia}
pot(T_i)=\frac{\sum_{i=0}^{\Lambda -1} t_i^2}{\Lambda}
\end{equation}

The final volume is dependent on the speakers amplification of the signal. Thus, what matters is the relative potency of a note in relation to the other ones around it, or the potency of a music section in relation to the rest. Differences in volume are measured in decibels, calculated directly from the amplitudes through energy or potency:

\begin{equation}\label{decibels}
V_{dB}=10log_{10}\frac{pot(T^{'}_i)}{pot(T_i)}
\end{equation}

The quantity $V_{dB}$ has the decibel unit ($dB$). 
For each $10dB$ it is associated a "doubled volume".
Handy references are the $10dB$ for each step in the intensity scale: \emph{pianissimo}, \emph{piano}, \emph{mezzoforte}, \emph{forte} e \emph{fortissimo}. Other useful references are that equivalent in $dB$ of doubling amplitude or potency:

\begin{equation}\label{eq:ampVol}
\begin{split}
t_i^{'}=2 . t_i \Rightarrow pot(T^{'}_i)=4 . pot(T_i) \Rightarrow \\ \Rightarrow V^{'}_{dB}=10log_{10} 4 \approx 6 dB
\end{split}
\end{equation}

\begin{equation}\label{eq:potVol}
\begin{split}
pot(T^{'}_i)=2 pot(T_i) \Rightarrow \\ \Rightarrow V^{'}_{dB}=10log_{10} 2 \approx 3 dB
\end{split}
\end{equation}

and the amplitude gain for a sequence whose volume has been doubled ($10dB$) is:

\begin{equation}\label{eq:dobraVol}
\begin{split}
10log_{10}\frac{pot(T^{'}_i)}{pot(T_i)} = 10 \quad \Rightarrow \\ \Rightarrow \quad \sum_{i=0}^{\lfloor \Delta.f_a \rfloor -1}t^{'2}_i=10\sum_{i=0}^{\Lambda-1}t_i^2=\sum_{i=0}^{\Lambda-1}(\sqrt{10}.t_i)^2 \\
\therefore \quad t^{'}_i=\sqrt{10}t_i \quad \Rightarrow \quad t^{'}_i \approx 3,16t_i
\end{split}
\end{equation}

Thus, it is necessary a little bit more than triplicate the amplitude for a doubled volume. These values are guides for increasing or decreasing the absolute values on the sample sequences with musical purposes. The conversion from decibels to amplitude gain (or attenuation) is straightforward:

\begin{equation}\label{ampDec}
A = 10^{\frac{V_{dB}}{20}}
\end{equation}

where $A$ is the multiplicative factor that relates the amplitudes before and after the amplification.

\subsection{Pitch}

As explained in the previous subsections, the musical particle (note) is a sequence $T_i$ in which duration and volume corresponds to the size of sequence and the amplitude of its samples. The pitch is specified by the fundamental frequency $f_0$ whose cycle has duration $\delta_{f_0}=1/f_0$. This duration, multiplied by the sampling frequency $f_s$ results on the number of samples per cycle: $\lambda_{f_0}=f_a . \delta_{f_0} =f_a/f_0$.

For didactic reasons, be $f_0$ such that it divides $f_s$ and $\lambda_{f_0}$ results an integer. If $T_i^f$ is a sonic sequence of fundamental frequency $f$, then:

\begin{equation}\label{periodicidade}
     T^f_i=\left\{ t_i^f \right\}=\left\{ t^f_{i+\lambda_{f}}  \right\}= \left\{ t^f_{i+\frac{f_a}{f}} \right\}
\end{equation}

In the next section, frequencies $f$ that does not divide $f_s$ will be considered. We will see that this restriction does not imply in lost of generality of this section's content.

\subsection{Timbre}

While the wave period corresponds to the fundamental frequency, the trajectory of the wave inside the period - called the waveform - defines a harmonic spectrum and, thus, a timbre\footnote{Timbre is a subjective and complex characteristic. Physically, the timbre is multidimensional and given by the temporal dynamics behavior of energy in the spectral components that are harmonic or noisy. Beyond that, the word timbre is used to designate different things: one same note has different timbres, a same instrument has different timbres, two instruments of the same family have, at the same time, the same timbre that blends them in the same family, and different timbres as they are different instruments. It is worth to mention that timbre is not always manifested in spectral traces, since cultural or circumstantial aspects alter our perception of timbre}. Musically, it matters that sonic spectra with minimum differences can result in timbres with crucial expressive differences and, consequently, different timbres can be produced by using different spectra\cite{Roederer}.

The simplest (and most important case, as we will see in the following) is the spectrum that consists of only of the fundamental frequency $f$. This defines the sinusoid: a frequency in pure oscillatory movement called 'simple harmonic movement'. Let $S_i^f$ be a sequence whose samples $s_i^f$ describes a sinusoid of frequency $f$:

\begin{equation}\label{senoide}
     S^f_i=\{ s^f_i \}=\Bigl\{ \sin\bigl(2\pi \frac{i}{\lambda_f} \bigr)  \Bigr\} = \Bigl\{ \sin\bigl(2\pi f \frac{i}{f_a}\bigr)  \Bigr\} 
\end{equation}

where $\lambda_f=\frac{f_a}{f}=\frac{\delta_f}{\lambda_a}$  is the number of samples in the period.

In a similar fashion, other waveforms are applied in music considering their spectral qualities and simplicity. While the sinusoid is an isolated point in the spectrum, these additional waves present a succession of harmonic components of the sinusoid. These waveforms are specified in equations~\ref{sinusoid},~\ref{sawTooth},~\ref{triangular} and~\ref{square} and illustrated in Figure~\ref{fig:formasDeOnda}.
Although the artificial waveforms are traditionally used in music for synthesis and oscillatory control of variables, they are also useful outside music context\cite{Openheim}.

The sawtooth presents all components of the harmonic series with decreasing energy of $-6dB/octave$. The sequence of temporal samples can be described as following:

\begin{equation}\label{sawTooth}
     D^f_i=\left\{ d^f_i \right\}=\left\{ 2\frac{i\,\%\lambda_f}{\lambda_f} -1 \right\}
\end{equation}

The triangular waveform present only odd harmonics falling with $-12dB/octave$:
\begin{equation}\label{triangular}
     T^f_i=\left\{ t^f_i \right\}=\left\{1- \left| 2 - 4\frac{i\,\%\lambda_f}{\lambda_f} \right| \right\}
\end{equation}

The sawtooth is a common starting point for a subtractive synthesis, because it has both odd and even harmonics with high energy. For musical intention, these waveforms are excessively rich in sharp harmonics, and attenuator filtering on treble and middle parts of the spectrum is specially useful for reaching a more natural and pleasant sound. 
The relatively attenuated harmonics of the triangle wave makes it the more functional - among the listed ones - to be used in the synthesis of musical notes without any treatment.

The square wave preservers only odd harmonics falling at $-6dB/octave$:

\begin{equation}\label{quadrada}
     Q^f_i=\left\{ q^f_i \right\}= \left\{
         \begin{array}{l l}
              1 & \quad \text{for } \; \; (i\,\%\lambda_f)   <  \lambda_f /2  \\
              -1 & \quad \text{otherwise}\\
         \end{array} \right.
\end{equation}

The square wave can be used in a subtractive synthesis with the purposes of mimicking a clarinet. This instrument has only the odd harmonic components and the square wave is convenient with its abundant energy in high frequencies.

\begin{figure*}
    \centering
        \includegraphics[width=\columnwidth]{figures/waveForms}
    \caption{Basic musical waveforms: (a) the synthetic  waveforms; (b) the real waveforms.}
        \label{fig:formasDeOnda}
\end{figure*}

Figure~\ref{fig:formasDeOnda} presents the waveforms described in equations  ~\ref{senoide}, ~\ref{denteDeSerra}, ~\ref{triangular} and ~\ref{quadrada} for $\lambda_f=100$ (period of $100$ samples). If $f_s=44.1kHz$, a PCM standard in Compact Disks, the wave has fundamental frequency $f=\frac{f_a}{\lambda_f}=\frac{44100}{100} = 441 \; Herz$, reflecting the A4, just above the central "C", whatever the waveform is.

The spectrum of each basic waveform is in Figure~\ref{fig:espectroDeondas}. The isolated and exactly harmonic components of the spectrum is a consequence of the fixed period usage. The sinusoid consists of a one and only node in spectrum (**esquisito o texto**), pure frequency. As we said before, the sawtooth is the only waveform with a complete harmonic series (odd and even components). Triangular and square waves has the same components (odd harmonics), decaying at $-12dB/octave$ and $-6dB/octave$, respectively.

\begin{figure*}
    \centering
        \includegraphics[width=\columnwidth]{figures/waveSpectrum}
    \caption{Spectrum of basic artificial musical waveforms.}
        \label{fig:espectroDeOndas}
\end{figure*}

The harmonic spectrum is formed by frequencies that are multiple from the fundamental frequency $f_n=(n+1)f_0$. As the human linear perception of pitch follows  a geometric progression of frequencies, the spectrum has notes different from the fundamental frequency. Additionally, the number of harmonics will be limited to the maximum frequency $f_s/2$ (by Nyquist's theorem).

From a musical perspective, to internalize that energy in a component of frequency $f_n$ means an oscillation in the constitution of the sound, purely harmonic and in that frequency $f_n$. This energy, specifically concentrated on the $f_n$, is separated by the ear to enter in a cognitive level of processing (this separation is done in many species with mechanisms similar to the human cochlea~\cite{Roederer}).

The sinusoidal components are usually the main responsible for the timbre quality. If they are not presented in harmonic proportions (small number relations), the sound is perceived as noisy or dissonant, in opposite of a sonority with a unequivocally established fundamental frequency when they are present. Furthermore, the notion of absolute pitch in a complex sonority is based relies on the similarity of the spectrum to the harmonic series~\cite{Roederer} (**nao entendi o q vcs querem dizer**).

In the case of a fixed length period and waveform, the spectrum is perfectly harmonic and static. Each waveform is compound of specific proportions of harmonic components and the greater the curvature of the part, the greater the contribution of the part to the energy on the high harmonics. This aspect can be seen on real sounds. The wave labeled as ``sampled real sound'' in Figure~\ref{fig:formasDeOnda} has a period of $\Lambda_f=114$ samples, extracted from a relatively well behaved real sound. The oboe wave was sampled for A4 also in $44.1kHz$. The chose period for sampling was a relatively short one, with $98$ samples, and corresponds to the frequency $\frac{44100}{98}=450Hz$. It can be noticed from the curvatures, the oboe's rich spectrum in high frequencies and the lower spectrum of the real sound.

The sequence 
$ R_i=\{ r_i \}_0^{\lambda_f-1}$ of samples in the real sound of Figure~\ref{fig:formasDeOnda} can be taken as basis for a sound $T_i^f$ in the following way:

\begin{equation}\label{sampleandoFormaDeOnda}
     T^f_i=\{ t_i^f \}=\Bigl\{ r_{(i\,\%\lambda_{f})} \Bigr\}
\end{equation}

The resulting sound has the momentary spectrum of the original sound. Consequence of its repetition in an identical form, the spectrum is perfectly harmonic, without noise and with variations typical of the natural phenomenon. This can be observed in Figure~\ref{fig:espectroOboe}, that shows the spectrum of the original oboe note and a note with same duration and whose samples consists of the repetition of cycle of Figure~\ref{fig:formasDeOnda}. The natural spectrum exhibits variations in the frequencies of the harmonics, in their intensities and some noise. The note made from the sampled period has a perfectly harmonic spectrum.

\begin{figure*}
    \centering
        \includegraphics[width=\columnwidth]{figures/oboeNaturalSampledSpectrum}
    \caption{Spectrum of the sonic waves of a natural oboe note and one made from a sampled period. The natural sound has fluctuations in the harmonics and in its noise, while the sampled period note has a perfectly harmonic spectrum.}
        \label{fig:espectroOboe}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Spectrum at sampled sound}

The presence and behavior of these sinusoidal components in the discretized sound have some particularities. Considering a signal $T_i$ and its corresponding Fourier decomposition $\mathcal{F}\langle T_i\rangle=C_i=\{c_i\}_0^{\Lambda-1}$, the recomposition is the sum of the frequencies components as time samples\footnote{It is important to note that the factor $\frac{1}{\Lambda}$ could be distributed among the Fourier transform and its reconstruction, as preferred.}:

\begin{multline}\label{recomposicaoFourier}
t_i = \frac{1}{\Lambda}\sum_{k=0}^{\Lambda-1}c_ke^{j \frac{2\pi k}{\Lambda} i } \\ = \frac{1}{\Lambda}\sum_{k=0}^{\Lambda-1}(a_k+ j . b_k)\left[cos(w_k i)   +j . sen(w_k i)\right]
\end{multline}

where $c_k = a_k + j . b_k$ defines the amplitude and phase of each frequency: $w_k=\frac{2\pi}{\Lambda}k$ in radians or $f_k=w_k\frac{f_a}{2\pi}=\frac{f_a}{\Lambda}k$ in Hertz, taking into account the respective limits in $\pi$ and in $\frac{f_a}{2}$ given by the Nyquist Theorem.

Equation~\ref{recomposicaoFourier} shows how the imaginary term $c_k$ adds a phase to the real sinusoid. For example, the imaginary terms $b_k$ from the Fourier spectral decomposition make possible the phase sweep $\left[-\frac{\pi}{2},+\frac{\pi}{2}\right]$ given by $tg^{-1}\left(\frac{b_k}{a_k}\right)$ which has this image. The signal from $a_k$ specifies the right and left side of the trigonometric circle, complementing the phase sweeping: $\left[-\frac{\pi}{2},+\frac{\pi}{2}\right] \cup \left[\frac{\pi}{2},\frac{3\pi}{2}\right]\equiv [2\pi]$.

For a sound signal, samples $t_i$ are real and are given by the real part of equation~\ref{recomposicaoFourier}:

\begin{equation}\label{moduloEfase}
\begin{split}
t_i& = \frac{1}{\Lambda}\sum_{k=0}^{\Lambda-1}\left[a_k cos(w_k i) -b_k sen(w_k i)\right] \\
   & = \frac{1}{\Lambda}\sum_{k=0}^{\Lambda-1}\sqrt{a_k^2 + b_k^2} \; cos\left[w_k i - tg^{-1}\left(\frac{b_k}{a_k}\right)\right]
\end{split}
\end{equation}

 \begin{figure}[h!]
     \centering
         \includegraphics[width=\columnwidth]{figures/amostras2c__}
     \caption{Oscillation for 2 samples (maximum frequency for any $f_a$). The first coefficient reflects a detachment (\emph{offset} or \emph{bias}) and the second coefficient specifies the oscillation amplitude.}
         \label{fig:amostras2}
 \end{figure}

Figure~\ref{fig:amostras2} shows two samples and its spectral components. In this case, the Fourier decomposition has one unique pair of coefficients $\{c_k=a_k-j.b_k\}_0^{\Lambda-1=1}$ relatives to frequencies $\{f_k\}_0^1=\left\{w_k\frac{f_a}{2\pi}\right\}_0^1=\left\{k\frac{f_a}{\Lambda=2}\right\}_0^1=\left\{0,\frac{f_a}{2}=f_{\text{max}}\right\}$
with energies $e_k=\frac{(c_k)^2}{\Lambda=2}$. The role of amplitudes $a_k$ is clearly observed with $\frac{a_0}{2}$, the fixed detachment\footnote{Also called \emph{bias} or \emph{offset}.} and $\frac{a_1}{2}$, oscillation own amplitude given by the relation $f_k=k \frac{f_a}{\Lambda=2}$.
This case have special relevance. It is necessary the of 2 samples to represent an oscillation and it yields the Nyquist frequency $f_{\text{max}}=\frac{f_a}{2}$. That is the maximum frequency in a sound sampled with $f_a$ samples per second\footnote{Any sampled signal has this property, not only the digitalized sound.}.

All fixed sequences $T_i$ of only $3$ samples also have just $1$ frequency, since their first harmonic has $1,5$ samples and exceeds the bottom limit of 2 samples, e.g.\ the frequency of the harmonic would exceed the Nyquist frequency:  $\; \frac{2. f_a}{3} > \frac{f_a}{2} $. 
The coefficients $\{c_k\}_0^{\Lambda-1=2}$ are present in 3 frequency components. One is relative to zero frequency ($c_0$), and the other two ($c_1$ and $c_2$) have the same role for the reconstruction of sinusoid with $f=f_a/3$.

 \begin{figure}[h!]
     \centering
         \includegraphics[width=\columnwidth]{figures/amostras3b}
     \caption{Three fixed samples presents only one non-null frequency. $c_1=c_2^*$ and $w_1 \equiv w_2$.}
         \label{fig:amostras3}
 \end{figure}

$\Lambda$ real samples $t_i$ result in $\Lambda$ complex coefficients $c_k=a_k+j.b_k$. The coefficients $c_k$ are equivalent two by two, corresponding to the same frequencies and with same relevance\footnote{Equal real part and imaginary with inverse order: $a_{k1}=a_{k2}$ and $b_{k1}=-b_{k2}$. As consequence the modules are equal and phases have inverse order.}. Remembering that $f_k = k\frac{f_a}{\Lambda}, \; k \in \left\{0, ..., \left \lfloor \frac{\Lambda}{2} \right \rfloor \right\} $. When $k > \frac{\Lambda}{2}$, the frequency $f_k$ is mirrored by $\frac{f_a}{2}$ in this way: $f_k=\frac{f_a}{2} - (f_k-\frac{f_a}{2})=f_a-f_k=f_a - k\frac{f_a}{\Lambda}=(\Lambda-k)\frac{f_a}{\Lambda} \;\;\;\; \Rightarrow \;\;\;\; f_k\equiv f_{\Lambda-k} \; ,\;\; \forall \;\; k<\Lambda$.

The same could be observed in $w_k=f_k.\frac{2\pi}{f_a}$ and considering the periodicity $2\pi$, it results in $w_k=-w_{\Lambda-k}$. Given the cosine as an even function and the inverse tangent as an odd function, the components in $w_k$ and $w_{\Lambda-k}$ sums up in the equation that reconstructs the real samples, yielding equation~\ref{recomposicaoFourier}.

In other words, in a decomposition of $\Lambda$ samples, the $\Lambda$ given frequencies components $\{c_i\}_0^{\Lambda-1}$ are equivalents in pairs.
Except for $f_0$, and when having an even $\Lambda$, from $f_{\Lambda/2}=f_{\text{max}}=\frac{f_a}{2}$ both components are isolated, e.g.\ there is any other component in frequency $f_0$ or $f_{\Lambda/2}$ (if even $\Lambda$) regarding itself. 
It is true because $f_{\Lambda/2}=f_{(\Lambda-\Lambda/2) = \Lambda/2}$ and $f_0=f_{(\Lambda-0)=\Lambda}=f_0$.
Furthermore, these two frequencies (zero and maximum frequency) are not represented having phase variation, being strictly real. In this way, it is possible to conclude the number $\tau$ of equivalent coefficient pairs:

\begin{equation}\label{coefsPareados}
\tau = \frac{\Lambda - \Lambda \% 2}{2} +\Lambda \% 2 -1
\end{equation}

This make clear visible the equivalences ~\ref{equivalenciasFreqs}, ~\ref{equivalenciasModulos} and ~\ref{equivalenciasFases}:

\begin{equation}\label{equivalenciasFreqs}
f_{k}\equiv f_{\Lambda-k}\;, \;\; w_{k}\equiv-w_{\Lambda-k}\;\;\;, \quad \;\; \forall \quad 1 \leq k \leq \tau  
\end{equation}

\begin{figure}[h!]
    \centering
        \includegraphics[width=\columnwidth]{figures/amostras4__}
    \caption{Frequential components for 4 samples.}
        \label{fig:amostras4}
\end{figure}

Having $a_k = a_{\Lambda -k}\;\;$ and $\;\;b_k = - b_{\Lambda -k}$:

\begin{equation}\label{equivalenciasModulos}
\sqrt{a_k^2 + b_k^2} = \sqrt{a_{\Lambda - k}^2 + b_{\Lambda -k}^2} \;\;, \quad \;\; \forall \quad 1 \leq k \leq \tau  \\
\end{equation}

\begin{equation}\label{equivalenciasFases}
tg^{-1}\left(\frac{b_k}{a_k}\right)=-tg^{-1}\left(\frac{b_{\Lambda -k}}{a_{\Lambda - k}}\right)\;\;,\quad \;\; \forall \quad 1 \leq k \leq \tau
\end{equation}

with $k \in \mathbb{N}$.

The equation~\ref{moduloEfase} for the real signal reconstruction, the modules and phases equivalences~\ref{equivalenciasModulos} and~\ref{equivalenciasFases}, the number of paired coefficients~\ref{coefsPareados}, and equivalence of paired frequencies~\ref{equivalenciasFreqs} expose a general case for components combination in each sample $t_i$:

\begin{multline}\label{eq:reconsCompleta}
t_i = \frac{a_0}{\Lambda} + \frac{2}{\Lambda}\sum_{k=1}^{\tau}\sqrt{a_k^2 + b_k^2} \; cos\left[w_k i - tg^{-1}\left(\frac{b_k}{a_k}\right)\right]+ \\ \frac{ a_{\Lambda/2}}{\Lambda}.(1-\Lambda\% 2)
\end{multline}

\begin{figure}[h!]
    \centering
        \includegraphics[width=\columnwidth]{figures/amostras4formas__}
    \caption{Basic wave forms for 4 samples.}
        \label{fig:formas4}
\end{figure}

Therefore, as in Figure~\ref{fig:amostras3}, the Fourier transform of 3 samples have 2 frequencies coefficients with same amount of energy in the same frequency.

With 4 samples it is possible to represent 1 or 2 frequencies with any proportions. Figure~\ref{fig:amostras4} depicts wave forms for 4 samples and its respective two components. The individual contributions sums together yielding the original wave form and a brief inspection reveals the major curvatures resulting from the higher frequency, while a fixed detachment of component summation results from the zero frequency component.

\begin{figure}[h!]
    \centering
        \includegraphics[width=\columnwidth]{figures/amostras6}
    \caption{Frequencies components for 6 samples: 3 sinusoid sums with \emph{bias}.}
        \label{fig:amostras6}
\end{figure}

Figure~\ref{fig:formas4} shows the harmonics for 4 samples in basic wave forms of equations ~\ref{senoide}, ~\ref{denteDeSerra}, ~\ref{triangular} and ~\ref{quadrada}. Their conjunction results in only 1 sinusoid, with exception to sawtooth wave, which have even harmonics.

Figure~\ref{fig:amostras6} presents a sinusoidal decomposition for the 6 samples case and Figure~\ref{fig:formas6} presents the decomposition of the basic wave forms.
In this case all wave forms are different in spectrum: square and triangular ones have the same components but with different proportions, while the sawtooth have an extra component.

\begin{figure}[h!]
    \centering
        \includegraphics[width=\columnwidth]{figures/amostras6formas__}
    \caption{Basic wave forms for 6 samples: triangular and square wave forms have odd harmonics, with different proportions and phases; the sawtooth wave also has even harmonics.}
        \label{fig:formas6}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The basic note}\label{notaBasica}

Consider $f$ presented in a such way that divides $f_a$\footnote{As pointed before, this limitation simplify the explanation without losing generality, and will be overcome in the next section.}. A sequence $T_i$ of sound samples are separated by $\delta_a=1/f_a$ and describes a musical note with a frequency of $f$ Hertz and $\Delta$ seconds of duration if, and only if, it has the periodicity equal to $\lambda_f=f_a/f$ and size $\Lambda=\lfloor f_a . \Delta \rfloor $:

\begin{equation}\label{eq:notaBasica}
T_i^{f,\; \Delta}=\{t_{i \, \% \lambda_f} \}_0^{\Lambda-1}= \left \{t^f_{i \; \% \left( \frac{f_a}{f} \right) } \right \}_0^{\Lambda-1}
\end{equation}

The note by itself does not specifies a timbre. Nevertheless, it is necessary to choose the right waveform so that the samples $t_i$ have a value established individually. A unique period from the basic waveforms can be used to specify the note, where $\lambda_f=\frac{f_a}{f}$ is the number of samples at the period. Here, $L_i^{f,\, \delta_f} $ is the sequence that describes a period of the waveform $L_i^f \in \{S_i^f,Q_i^f,T_i^f,D_i^f,R_i^f \}$ with duration $\delta_f=1/f$ (as given by equations~\ref{senoide}, ~\ref{denteDeSerra}, ~\ref{triangular} and ~\ref{quadrada}) and $R_i^f$ is a sampled real waveform:

\begin{equation}\label{periodoUnico}
L_i^{f , \delta_f } = \left\{ l_i^f \right\}_0^{\delta_f . f_a -1}=\left\{ l_i^f \right\}_0^{\lambda_f-1}
\end{equation}

Therefore, the sequence $T_i$ will consist in a note of duration $\Delta$ and frequency $f$ if:

\begin{equation}\label{eq:notaBasicaTimbre}
T_i^{f,\; \Delta}=\left\{t_i^f\right\}_0^{\lfloor f_a . \Delta \rfloor -1}=\left \{ l^f_{i\,\%\left(\frac{f_a}{f}\right)} \right \}_0^{\Lambda-1}
\end{equation}

\subsection{Spatial localization and spatialization}\label{subsec:spac}

Although it is not one of its four basic properties, a musical note always has a spatial localization: the note source position at the ordinary physical space. In this context, the existence of an environment that reverbs the played note it is the matter of 'spatialization'. Both fields, spatialization and spatial localization, are widely valued by audiophiles and music industry~\cite{floEsp}.

\subsubsection{Spatial localization}

It is believed that the perception of spatial localization of sound occurs in our nervous system by three information: the delay of incoming sound between both ears, the difference of sound intensity at each ear and the filtering realized by the human body, including its chest, head and ears~\cite{Roederer, hrtf, Heeger}.

\begin{figure}[h!]
    \centering
        \includegraphics[width=.5\textwidth]{figures/espacializacao___}
    \caption{Detection of sound source spatial localization: schema used to calculate Interaural Time Difference (ITD) and Interaural Intensity Difference (IID).}
    \label{fig:spac}
\end{figure}

Considering only the direct incidences in each ear the equations are quite simple. Given the distance between ears $\zeta$\footnote{It is known that $\zeta \approx 21,5cm$ to an adult human}, an object placed at $(x,y)$ as in Figure~\ref{fig:spac} is distant of each ear by:

\begin{equation}\label{eq:distOuvidos}
\begin{split}
d & =\sqrt{\left (x-\frac{\zeta}{2} \right )^2+y^2} \\
d' & =\sqrt{\left (x+\frac{\zeta}{2} \right )^2 + y^2}
\end{split}
\end{equation}

\noindent and immediate calculations results in the Interaural Time Difference:

\begin{equation}\label{eq:dti}
ITD=\frac{d'-d}{v_{sound\;at\;air}\approx 343.2 }\quad \text{seconds}
\end{equation}

\noindent and in the Interaural Intensity Difference:

\begin{equation}\label{eq:dii}
IID=20\log_{10}\left (\frac{d}{d'}\right) \quad decibels
\end{equation}

Converting to amplitude it gives $IID_a=\frac{d}{d'}$. The $IID_a$ can be used as a multiplicative constant to the right channel of a stereo sound signal: $\{t_i'\}_0^{\Lambda -1}=\{IID_a . t_i\}_0^{\Lambda -1}$. It is possible to use IID together with ITD as a time advance for the right channel in relation with the left one. It is a crucial entail to localization in bass sounds and percussive sonorities~\cite{Heeger}. 
Considering $\Lambda_{ITD}=\lfloor ITD . f_a \rfloor$:

\begin{equation}\label{eq:locImpl}
\begin{split}
\Lambda_{ITD} & = \left \lfloor \frac{d'-d}{343,2}  f_a \right \rfloor \\
IID_a & = \frac{d}{d'} \\
\left\{t_{(i+\Lambda_{ITD})}'\right\}_{\Lambda_{ITD}}^{\Lambda+\Lambda_{ITD}-1} & =\left\{IID_a . t_i\right\}_0^{\Lambda-1} \\
\left\{t_i'\right\}_0^{\Lambda_{ITD}-1} & = 0
\end{split}
\end{equation}

\noindent with $t_i$ as the right channel and $t_i'$ the left channel. If $\Lambda_{ITD} < 0 $, it is only needed to change $t_i$ by $t_i'$ and to use $\Lambda_{ITD}'= | \Lambda_{ITD} | $.

Although simple until here, spatial localization depends considerably of other steps. Using ITD and IID it is possible to specify only the horizontal angle (azimuthal) $\theta$ given by:

\begin{equation}\label{eq:angulo}
\theta=\tan^{-1}\left ( \frac{y}{ x }  \right )
\end{equation}

\noindent with $x,y$ as presented in Figure~\ref{fig:spac}. Henceforth, there are problems when $\theta$ focuses in the so called "cone of confusion": where the same pair of ITD and IID results in a large number of points inside the code. On those points the inference of azimuthal angle depends specially of the attenuate filtering for high frequencies, since the head interferes much more in the treble mechanical waves than in the bass waves~\cite{Heeger,hrtf}. It is also relevant to the hearing of source side, where the sound has enough low frequencies there is a diffraction and the wave comes to the ear with a delay of $\approx 0,7ms$.\cite{floEsp}

Figure~\ref{fig:spac} also shows an acoustic shadow from cranium, an important phenomenon to perception of source azimuthal angle in the cone of confusion. The cone itself is not shown in Figure~\ref{fig:spac} because it is not exactly a cone and its precise dimensions were not encountered in the literature. This makes it hard to conceive given the filtering and diffraction dependent of the sound spectrum. In this way, the cone of confusion can be understood as a cone with its top side placed in the middle of head and growing out in the direction of each ear~\cite{hrtf}.

On the other hand, the complete localization, including height and distance of sound source, is given by the Head Related Transfer Function (HRTF)~\cite{hrtf}. There are well known open databases of HRTF as CIPIC and it is possible to apply those transfer functions in a sound signal by convolution (see equation~\ref{eq:conv})~\cite{CIPIC}. The human body considerably changes the filtering and there are techniques to generate HRTFs to be universally used, as expected~\cite{lazaSPA}.

\subsubsection{Spatialization}

Spatialization results from sound reflections and absorptions by room/environment surface where the note is played. The sound propagates throw the air with a speed of $\approx 343,2m/s$ and can be sent from a source without any directionality pattern. When a sound pulse encounters a surface there is reflection. In this reflection occurs: 1) the component inversion of propagation speed whose is normal to the surface, as 2) the energy absorption, specially in trebles. The sound waves propagate until they reach inaudible levels. The moment when some sound pulse reach the human ear can be described as the moment it reaches the ear and the absorption filters of each surface it has reached. It is possible to simulate reverberations that are impossible in real systems. To experimentation, it is possible to use asymmetric reflections with relation to the axis perpendicular to the surface, or to increase specific frequency bands (known as 'resonances'), both characteristics are now found in real systems.

There are some reverberation models that are less related to the calculation of each reflection, exploring valuable information of the auditory system. In fact, reverberation can be modeled with a set of 2 temporal and spectral characteristics:

\begin{itemize}
   \item First period: 'first reflections' are more intense and scattered.
   \item Second period: 'late reverberation' is practically a dense succession of indistinct delays with exponential decay and statistical occurrences.
   \item First band: bass has some resonance frequencies relatively spaced.
   \item Second band: mid and treble have progressive decay and smooth statistical fluctuations.
\end{itemize}

Smith III points out that reasonable concert rooms have total reverberation time of approx. $1,9$ seconds, and that the period of first reflections is around $0,1$ seconds. These values suggests that, in the given conditions, there are perceived wave pulses which propagates for $652,08$ meters long ($83,79k$ samples in $f_a=44,1kHz$) before reaching the ear. In addition, sound reflections made after propagation for $34,32$ meters long ($4,41k$ samples in $f_a=44,1kHz$) have a tangled form whose incidences are less distinct by hearing. These first reflections are particularly important to space sensation. The first incidence is the direct sound, described by ITD and IID of equations~\ref{eq:dti} and~\ref{eq:dii}. Admitting that each one of the first reflections, before reaching the ear, will propagate at least $3-30m$ dependents of room dimension, the separation between the first reflection is at  $8-90ms$ ($\approx 350-4000$ samples in $f_a=44.1kHz$). It is possible to experimentally verify that the number of reflections increases with squared proportion $\approx k.n^2$. A discussion about the use of convolutions and filtering to favor these implementations is provided in subsection~\ref{subsec:mus2}, specially at paragraphs about reverberation.

\subsection{Musical use}\label{subsec:basMus}

Once we have well defined the basic note, it is time to build musical structures with sequences based on those particles. The sum of elements with same index of $N$ sequences $T_{k,i}=\{t_{k,i}\}_{k=0}^{N-1}$ with same size $\Lambda$ results in overlapped spectral contents in a process called sound mixing:

\begin{equation}\label{eq:mixagem}
\{t_i\}_0^{\Lambda-1}=\left \{ \sum_{k=0}^{N-1}t_{k,i} \right \}_0^{\Lambda-1}
\end{equation}

\begin{figure}[h!]
    {\centering
        \includegraphics[width=\columnwidth]{figures/mixagem}}
    \caption{Mixing of three sound sequences. The amplitudes are directly overlapped.}
        \label{fig:mixagem}
\end{figure}

Figure~\ref{fig:mixagem} illustrates this overlapping process of discretized sound waves, disposing 100 samples. It is possible to conclude that, if $f_a=44.1kHz$, the frequencies of the sawtooth, square and sine wave are, respectively: $\frac{f_a}{100/2}=882Hz$, $\frac{f_a}{100/4}=1764Hz$ and $\frac{f_a}{100/5}=2205Hz$. The samples duration are very short $\frac{f_a=44.1kHz}{100} \approx 2 \text{ milliseconds}$. Someone just need to complete the sequence with zeros to sum up sequences with different sizes.

The mixed notes are generally separated by the ear according to the physic laws of resonance and by the nervous system~\cite{Roederer}.  This musical notes mixing process results in musical harmony whose intervals between frequencies and chords of simultaneous notes guide subjective and abstract aspects of music and its appreciation~\cite{Harmonia}.

The sequences can be concatenated in time. If the sequences $\{t_{k,i}\}_0^{\Lambda_k-1}$ of size $\Lambda_k$ represent $k$ musical notes, their concatenation in a unique sequence $T_i$ is a simple musical sequence or melody:

\begin{multline}\label{eq:concatenacao}
\{t_i\}_0^{\sum\Delta_k-1}=\{t_{l,i}\}_0^{\sum\Delta_k-1}, \;\; \\ l\text{ smallest integer } : \quad \Lambda_l > i -\sum_{j=0}^{l-1}\Lambda_j
\end{multline}

This mechanism is demonstrated in Figure~\ref{fig:concatenacao} with same sequences of Figure~\ref{fig:mixagem}. Although the sequences are short for the used sample rates, it is possible to observe the concatenation of sound sequences. Beside that, each note has it duration larger than $100ms$ if $f_a<1kHz$.

\begin{figure}[h!]
{    \centering
        \includegraphics[width=\columnwidth]{figures/concatenacao}}
    \caption{Concatenation of three sound sequences by temporal overlap of its samples.}
        \label{fig:concatenacao}
\end{figure}

The musical piece \emph{reduced-fi} explores the temporal overlapping of notes, resulting in a homophobic piece. The vertical principle is demonstrated at the \emph{sound boards}, static sounds with peculiar spectrum. Both pieces were written in Python at Appendices~\ref{ap:quadros} and~\ref{ap:reduced}, and are available as part of the \massa \emph{toolbox}.\cite{MASSA}

Now that the basic digital music note is described, the next section develops the temporal evolution of its contents as in \emph{glissandi} and volume envelopes. The filtering of spectral components and noise generation complements the definition of the musical note as a self contained unity. Section~\ref{notasMusica} is dedicated to the organization of these notes as music by using metrics and trajectories.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Variation in the basic musical note}\label{sec:internalVar}

The basic digital music note defined in section~\ref{sec:notaDisc} has following parameters: duration, pitch, intensity (volume) and timbre. This is a useful and paradigmatic modeling but does not covers all the aspects regarding a musical note.

At first place, the characteristics of the note are modified along the duration of the note itself~\cite{Chowning}. For example, a note played in a piano with duration of 3 seconds has abrupt intensity at its beginning time of execution. Then, it behaves with progressively decays, spectrum variations, presenting harmonics decaying before others and presenting harmonics showing up along the time.
Those variations are not mandatory but they are guidance from sound synthesis to musical uses, because it suitable reflects how sounds are represented in nature\footnote{The gold tip here is: to make a sound that incites interest by itself, do internal variations on it~\cite{Roederer}.}.
To explore all the possible ways which those variations occur is out of the scope of this study, given the considerably sensibility of the human ear and the complexity of our sound cognition. In the following we point out primary resources to produce those characteristic variations in the basic note. It is worthwhile to mention that all the described relations in the following section are currently implemented in Python at Appendix~\ref{sec:cod2}. The music pieces \emph{Transita para metro}, \emph{Vibra e treme}, \emph{Tremolos, vibratos e a frequência}, \emph{Trenzinho de caipiras impulsivos}, \emph{Ruidosa faixa}, \emph{Bela rugosi}, \emph{Chorus infantil}, \emph{ADa e SaRa} are listed in Appendices~\ref{ap:transita}, \ref{ap:vibra}, \ref{ap:tremolos}, \ref{ap:trenzinho}, \ref{ap:ruidosa}, \ref{ap:bela}, \ref{ap:chorus} and \ref{ap:ada}. The code is also part of the toolbox \massa, available online.\cite{MASSA}

\subsection{Lookup Table}\label{subsec:lookup}

The \emph{Lookup Table} (or simply LUT), is a data structure array which substitutes continuous and repetitive calculation by indexed access. It is frequently used to reduce computational complexity and to make possible the use of functions without the possibility of direct calculation, as the data sampled from nature.
In music its usage transcends the former, simplifying operations and making possible to use a single wave period to synthesize sounds in the whole audible spectrum, no matter what kind of wave was originally sampled.

\begin{figure}[h!]
    \centering
        \includegraphics[width=\columnwidth]{figures/lut}
    \caption{Search procedure in an indexed array (known as \emph{Lookup Table}) to synthesize sounds in different frequency using an unique waveform with high resolution.}
        \label{fig:lut}                                                                                                            
\end{figure}

Being $\widetilde{\Lambda}$ the wave period size and $\widetilde{L_i} = \left\{\, \widetilde{l}_i \,\right\}_0^{\widetilde{\Lambda} -1}$ the elements $\widetilde{l_i}$ of any wave period (refer to Equation~\ref{periodoUnico}), a sequence $T_i^{f,\,\Delta}$ with samples of a sound with frequency $f$ and duration $\Delta$ can be obtained by means of $\widetilde{L_i}$ as:

\begin{multline}\label{eq:lut}
T_i^{f,\,\Delta}=\left\{t_i^f\right\}_0^{\lfloor \, f_a . \Delta \, \rfloor -1} = \left\{ \, \widetilde{l}_{\gamma_i \% \widetilde{\Lambda} }\, \right\}_{0}^{\Lambda-1}\; , \quad \\ \text{where} \;\; \gamma_i = \left \lfloor i . f \frac{ \widetilde{\Lambda}}{f_a} \right \rfloor  
\end{multline}

In other words, with the right LUT indexes ($\gamma_i\%\widetilde{\Lambda}$) it is possible to synthesize a sound with any frequency. Figure~\ref{fig:lut} illustrates the calculation of a $\{t_i\}$ sample from $\left\{\,\widetilde{l}_i\,\right\}$ for a frequency of $f=200Hz$, $\widetilde{\Lambda}=128$ and considering the sample rate as $f_a=44.1kHz$.
Though this is not a practical configuration (as assigned below), it makes possible a graphical visualization of the procedure.

The calculation of the integer $\gamma_i$ introduces noise which decreases as $\widetilde{\Lambda}$ increases.
In order to use this calculation in sound synthesis, with $f_a=44.1 kHz$ the standard guidelines suggest the use of $\widetilde{\Lambda} = 1024$ samples, since it does not produce relevant noise considering the audible spectrum. The round or interpolation method are not decisive in this process~\cite{Geiger}.

The expression defining the variable $\gamma_i$ can be understood as $f_a$ being summed to $i$ at each $1$ second.
If $i$ is divided by the sample frequency, it results in $\frac{i}{f_a}$,
which is incremented in $1$ at each $1$ second. Multiplied by the period length it results in $i \frac{\widetilde{\Lambda}}{f_a}$ that look up the period in $1$ second. Finally, with frequency $f$ it results in $i . f \frac{\widetilde{\Lambda}}{f_a}$ which complements $f$ lookups of the period $\widetilde{\Lambda}$ in $1$ second, i.e., the resulting sequence presents the fundamental frequency $f$.

There are important considerations here: it is possible to use any value for $f$, the limits exist only in the loud frequencies when the size of table  $\widetilde{\Lambda}$ is not sufficient enough considering the sample rate $f_a$. The lookup table procedure is computationally costless and replaces the calculations by simple indexed searches (what is generally understood as an optimization process). Unless otherwise stated, this procedure will be used along all the text for every applicable case.

LUTs are broadly used in computational implementations for music. A classical usage of LUTs is known as \emph{Wavetable Synthesis}, which consists of many LUTs used together to generate a music note quasi-periodic~\cite{Cook,Wavetable}.

\subsection{Incremental Variations of Frequency and Intensity}\label{subsec:vars}

Following the Weber and Fechner~\cite{Weber-Fechner} law, the human perception has a logarithmic relation with the causing stimulus. In other words, a stimulus following a exponential progression is perceived as linear.
For didactic reasons and given its use in AM and FM synthesis (refer to subsection~\ref{subsec:tvaf}), the linear variation is discussed first.

Considering the note with duration $\Delta = \frac{\Lambda}{f_a}$, the frequency $f=f_i$ varies linearly from $f_0$ to $f_{\Lambda -1}$. In this way, it is possible to define the following:

\begin{equation}\label{freqLinear}
 F_i=\{f_i\}_0^{\Lambda-1}=\left\{f_0 + (f_{\Lambda-1}-f_0)\frac{i}{\Lambda-1} \right\}_0^{\Lambda-1}
\end{equation}

\begin{multline}\label{indiceLinear}
 \Delta_{\gamma_i}=f_i\frac{\widetilde{\Lambda}}{f_a} \quad \Rightarrow \quad \gamma_i= \\ \left \lfloor \sum_{j=0}^{i} f_j\frac{\widetilde{\Lambda}}{f_a} \right \rfloor   =\left \lfloor \sum_{j=0}^{i} \frac{\widetilde{\Lambda}}{f_a} \left [f_0 + (f_{\Lambda-1}-f_0)\frac{j}{\Lambda-1} \right ] \right \rfloor 
\end{multline}

\begin{equation}\label{serieAmostralLin}
 \left\{t_i^{\;\overline{f_0,\, f_{\Lambda-1}}}\right\}_0^{\Lambda-1}=\left\{\,\widetilde{l}_{\gamma_i \% \widetilde{\Lambda}}\,\right\}_0^{\Lambda-1}
\end{equation}

where $\Delta_{\gamma_i}=f_i\frac{\widetilde{\Lambda}}{f_a}$ is the LUT increment between two samples given the sound frequency in the first sample.

Thus, it is possible to calculate the elements $t_i^{\;\overline{f_0,f_{\Lambda-1}}}$ having the period $\left\{\widetilde{l}_i\right\}_0^{\Lambda-1}$.

The equations~\ref{freqLinear}, \ref{indiceLinear} and \ref{serieAmostralLin} are related with the linear progression of the frequency. We recall that, for the general case, the frequency progression \emph{perceived} as linear follows an exponential progression\footnote{In other words, a geometric progression of frequency is perceived as an arithmetic progression of pitch.}.
It is possible to write this such that: $f_i=f_0 . 2^{\frac{i}{\Lambda-1} n_8}$ where  $n_8=\log_2\frac{f_{\Lambda-1}}{f_0}$ is the number of octaves between $f_0$ and $f_{\Lambda-1}$.
Therefore, $f_i=f_0 . 2^{\frac{i}{\Lambda-1}\log_2\frac{f_{\Lambda-1}}{f_0}}=
 f_0 . 2^{\log_2\left ( \frac{f_{\Lambda-1}}{f_0} \right )^{\frac{i}{\Lambda-1}}}=
 f_0 \left ( \frac{f_{\Lambda-1}}{f_0} \right ) ^{\frac{i}{\Lambda -1}}$. Accordingly, the transition equations of linear frequency for the human ear are given by:

\begin{multline}\label{freqExponencial}
 F_i=\{f_i\}_0^{\Lambda-1}= \\ \left\{f_0 \left ( \frac{f_{\Lambda-1}}{f_0} \right ) ^{\frac{i}{\Lambda -1}} \right\}_0^{\Lambda-1}
\end{multline}

\begin{multline}\label{indiceExponencial}
 \Delta_{\gamma_i}=f_i\frac{\widetilde{\Lambda}}{f_a} \quad \Rightarrow \\ \quad \gamma_i=\left \lfloor \sum_{j=0}^{i} f_j\frac{\widetilde{\Lambda}}{f_a} \right \rfloor   =\left \lfloor \sum_{j=0}^{i} f_0 \frac{\widetilde{\Lambda}}{f_a} \left ( \frac{f_{\Lambda-1}}{f_0} \right ) ^{\frac{j}{\Lambda -1}} \right \rfloor
\end{multline}

\begin{equation}\label{serieAmostralLog}
 \left\{t_i^{\;\overline{f_0,\,f_{\Lambda-1}}}\right\}_0^{\Lambda-1}=\left\{\,\widetilde{l}_{\gamma_i \% \widetilde{\Lambda}}\,\right\}_0^{\Lambda-1}
\end{equation}

\begin{figure}[h!]
     \centering
         \includegraphics[width=\columnwidth]{figures/transicao}
     \caption{Intensity transitions to different values of $\alpha$ (see equations~\ref{seqAmp} and ~\ref{transAmp}).}
         \label{fig:transicao}
\end{figure}

The term $\frac{i}{\Lambda-1}$ looks up the interval $[0,1]$ and it is possible to power it to a value in such a way that the beginning of the transition will be smoother or steeper.
This procedure is useful for energy variations of a vibratory wave to change the volume intensity\footnote{The volume change (psychophisical quality) is consequence of different sound characteristics like reverberation and concentration of treble harmonics, among which is wave energy. Wave energy is the easiest one to modify (see Equation~\ref{eq:potencia}) and it can vary in many ways. The most simple way consists of modifying the amplitude by multiplying the whole sequence by a real number. The increased energy without amplitude variation is the \emph{sound compression}, useful nowadays for music production~\cite{guillaume}}. It is sufficient to multiply the original sequence (being it generated or predetermined) by the sequence $a_{\Lambda-1}^{\left( \frac{i}{\Lambda-1} \right )^\alpha}$, where $\alpha$ is the given coefficient and $a_{\Lambda-1}$ is a fraction of the original amplitude, which is the aimed value to change until the end of transition.

Thus, for amplitude variations:

\begin{multline}\label{seqAmp}
 \{a_i\}_0^{\Lambda-1}=\left \{ a_0 \left ( \frac{a_{\Lambda-1}}{a_0} \right )^{\left ( \frac{i}{\Lambda-1} \right )^\alpha} \right \}_0^{\Lambda-1}= \\ \left \{ \left ( {a_{\Lambda-1}} \right )^{\left ( \frac{i}{\Lambda-1} \right )^\alpha} \right \}_0^{\Lambda-1} \text{ com } a_0=1
\end{multline}

\begin{equation}\label{transAmp}
 T_i^{'}=T_i \odot A_i = \{t_i . a_i\}_0^{\Lambda-1}=\left \{ t_i . (a_{\Lambda-1} )^{\left ( \frac{i}{\Lambda-1} \right )^\alpha} \right \}_0^{\Lambda-1}
\end{equation}

It is possible to take $a_0=1$ in order to start a new sequence with the original amplitude and then keep changing it with the following samples.
This restriction makes $a_{\Lambda-1}$ the amplitude variation term.
If $\alpha=1$, the amplitude variation follows the geometric progression whose defines the linear progression. Figure~\ref{fig:transicao} shows the transitions between values 1 and 2 and for different values of $\alpha$, a gain of $\approx 6dB$ as given by equation~\ref{eq:ampVol}.

An special attention should be driven while considering $a=0$.
In equation~\ref{seqAmp}, $a_0=0$ results in a division by zero and if $a_{\Lambda-1}=0$, there will be multiplication by zero. Both cases make the procedure useless, once no number different of zero can be represented as a ratio in relation of zero. It is possible to solve this dilemma choosing a number that is small enough like $-80dB\;\Rightarrow a=10^{\frac{-80}{20}}=10^{-4}$ as the minimum volume for a \emph{fade in} ($a_0=10^{-4}$) or for a \emph{fade out} ($a_{\Lambda-1}=10^{-4}$).

For linear amplification -- but not linear to perception -- it is sufficient to use an appropriate sequence $\{a_i\}$:

\begin{equation}\label{seqAmpLin}
a_i=a_0 + (a_{\Lambda-1}-a_0)\frac{i}{\Lambda-1}
\end{equation}

Here the conversation between decibels and amplitude is convenient. In this way, the equations~\ref{ampDec} and \ref{transAmp} specify a transition of $V_{dB}$ decibels:

\begin{equation}\label{seqAmpDB}
T_i^{'}=\left\{ t_i 10^{\frac{V_{dB}}{20}\left( \frac{i}{\Lambda-1} \right)^\alpha} \right\}_0^{\Lambda-1}
\end{equation}

\noindent for the general case of amplitude variations following a geometric progression. The greater the value of $\alpha$, the smoother the sound introduction and more intense its end. $\alpha>1$ results in volume transitions commonly called \emph{slow fade}, while $\alpha<1$ results in \emph{fast fade}.\cite{guillaume}

The linear transitions will be used for AM and FM synthesis and for applications of logarithmic transitions for tremolos and vibratos. A non-oscillatory exploration of those variations is in the music piece \emph{Transita para metro}, which code is in Appendix~\ref{ap:transita} and online at \massa.\cite{MASSA}

\subsection{Application of Digital Filters}\label{subsec:filtros}

This subsection is limited to a description of the procedure in sequences by convolution and differential equations, and the immediate applications. Its complexity exceeds to the scope of this study\footnote{The implementation of filters comprehends an area of recognized complexity with dedicated literature and software implementations. We recommended the reader to consult the bibliography~\cite{Openheim,smith}.}. The filters application can be part of the synthesis process or made individually as part of processes commonly called sound treatment.

\begin{itemize}
  \item  Convolution and finite impulse response (FIR) filters

\begin{figure}[h!]
     \centering
         \includegraphics[width=\columnwidth]{figures/convolucao}
     \caption{Graphical interpretation of convolution. Each resulting sample is the sum of the previous samples of a signal, with each one multiplied by the retrograde sequence of samples from the other signal.}
         \label{fig:conv}
\end{figure}

Filters applied by means of convolution are known by acronym FIR (Finite Impulse Response) and are characterized by having a time finite sample representation. This sample representation is called 'impulse response' $\{h_i\}$. The FIR filters are applied to the time domain of the digital sound by means of convolution with the respective impulse response of the filter\footnote{It is possible to apply the filter in the frequency domain multiplying the Fourier coefficients of both sound and the impulse response, and then performing the inverse Fourier transform in the resulting spectrum.\cite{Openheim}}. For this study, the convolution is defined as:

\begin{multline}\label{eq:conv}
 \begin{split}
 \left\{t_i'\right\}_0^{\Lambda_t+\Lambda_h-2\; = \;\Lambda_{t\, '}-1} =\{(T_j*H_j)_i\}_0^{\Lambda_{t \, '}-1} & = \\ \left \{ \sum_{j=0}^{min(\Lambda_h-1,i)}h_{j} . t_{i-j} \right \}_0^{\Lambda_{t\, '}-1} 
     & = \\ \left \{ \sum_{j=max(i+1-\Lambda_h,0)}^{i}t_j . h_{i-j} \right \}_0^{\Lambda_{t\, '}-1}
 \end{split}
\end{multline}

\noindent where $t_i=0$ for the samples not given beforehand.
In other words, the sound $\{t_i'\}$ resulting from the convolution of $\{t_i\}$ with the given impulse response $\{h_i\}$ has each $i$-th sample $t_i$ overwritten by the sum of its last $\Lambda_h$ samples $\{t_{i-j}\}_{j=0}^{\Lambda_h-1}$ multiplied one-by-one by samples of the another impulse response $\{h_i\}_0^{\Lambda_h-1}$. This procedure is illustrated in Figure~\ref{fig:conv}, where the impulse response $\{h_i\}$ is traveled by its retrograde form, and $t_{12}'$ and $t_{32}'$ are two calculated samples using the convolution given by $(T_j*H_j)_i=t_i'$. The resulting signal always has the length of $\Lambda_t+\Lambda_h -1=\Lambda_{t'}$.

With this procedure it is possible to apply reverbs, equalizers, \emph{delays} to name a few of a variety of other filters for sound processing, as well as musical/artistic effects.

The impulse response can be provided by physical measures or by pure synthesis. An impulse response for reverb application can be obtained by sound recording of the environment when someone triggers a snap which resembles an impulse or a sinusoidal lookup whose Fourier transform approximates its frequency response. Both are impulse responses which, properly convoluted with the sound sequence, result in the own sequence with a reverberation that resembles the physical one in the original environment where the measure happened~\cite{Cook}.

The inverse Fourier transform of an envelope even and real is equivalent to a impulse response of a FIR filter. It performs a frequency filtering with the envelope. The greater the number of samples, the higher the envelope resolution and the computational complexity as well. The convolution is recognized as computationally expensive.

An important property is the time shift caused by convolution with a shifted impulse. Despite computationally expensive, it is possible to create \emph{delay lines} by means of sound convolution with an impulse response that has an impulse to each relapse of sound.
In Figure~\ref{fig:delays} it is possible to observe the shifting caused by convolution with each impulse. Depending on the intensity of the impulses, the result is perceived in rhythm (around 20 impulses by second) or in pitch (around 20-40 impulses by second). In the last case, the processes resemble granular synthesis, delays, reverbs and equalization.

\begin{figure}[h!]
    \centering
        \includegraphics[width=\columnwidth]{figures/delays}
    \caption{Convolution with impulse: shifting (a), delay lines (b) and granular synthesis~(c). Disposed in increasing order of its pulse density.}
        \label{fig:delays}
\end{figure}

\item Infinite impulse response (IIR) filters

This class of filters is known by the acronym IIR. They are characterized by having an infinite time representation, i.e.\ the impulse response does not converges to zero. Its application is usually made by the following equation:

\begin{equation}\label{eq:diferencas}
 t_i' = \frac{1}{b_0}\left ( \sum_{j=0}^Ja_j . t_{i-j} + \sum_{k=1}^Kb_k . t_{i-k}' \right )
\end{equation}

\noindent with $b_0=1$ in the most cases it is possible to normalize the variables: $a_j'=\frac{a_j}{b_0}$ and $b_k'=\frac{b_k}{b_0} \Rightarrow b_0' = 1$.
Equation~\ref{eq:diferencas} is called 'difference equation' because the resulting samples $\left\{t_i'\right\}$ are given by differences between the original samples $\{t_i\}$ and the previous resulting ones $\left\{t_{i-k}'\right\}$.

There are many methods and tools to prepare IIR filters. The following texts list a selection of them for didactic purpose and to make easy future consulting. They are well behaved filters whose aspects are described in Figure~\ref{fig:iir}.

Considering filters with simple order, the cutoff frequency $f_c$ is the position where the filter performs an attenuation of $-3dB \approx 0.707 $ in the original amplitude.
For band-pass and band-reject (or 'notch') filters, this attenuation results on two specifications: $f_c$ (in this case, referred as 'center frequency') band width $bw$,
in both frequencies $f_c \pm bw$ there is an attenuation of $\approx 0.707$ in the original amplitude.
There is sound amplification in band-pass and band-reject filters when the cutoff frequency is low and the band width is large enough. In trebles those filters present only a shift in the expected waveform, expanding the envelope to the bass side of the band.

For filters whose frequency responses have other envelopes (for the module), it is possible to create chains and apply them successively. Another possibility is to use some biquad 'filter receipt'\footnote{Short for 'biquadratic': its transfer function has two poles and two zeros, i.e. its first direct form consists of two quadratic polynomials forming the fraction: $\mathbb{H}(z)=\frac{a_0+a_1.z^{-1}+a_2.x^{-2}}{1- b_1.z^{-1} -b_2 . z^{-2}}$.} or procedures for the calculation of coefficients of Chebichev filters\footnote{Butterworth and Elliptical filters can be considered as specific cases of Chebichev filters.\cite{Openheim,smith}}.
Both possibilities are explored by the referenced studies, in special~\cite{JOSFM,smith}, and the collection of filters maintained by the \emph{Music-DSP} community from University of Columbia.\cite{music-dsp,Openheim}

\end{itemize}

\begin{enumerate}
  \item Low-pass with simple pole and module of the frequency response in the upper left corner of Figure~\ref{fig:iir}. The general equation has as reference by the cutoff frequency $f_c \in (0,\frac{1}{2})$, fraction of the sample frequency $f_a$ that there is approximately an attenuation of $3dB$.
  The coefficients $a_0$ and $b_1$ of the IIR filter are given by means of the intermediate variable $x \in [e^{-\pi},1]$:

\begin{figure}[h!]
    \centering
        \includegraphics[width=\columnwidth]{figures/iir}
    \caption{Modules for the frequency response (a), (b), (c) and (d) for IIR filters of equations~\ref{eq:passa-baixas}, \ref{eq:passa-altas}, \ref{eq:passa-banda} and \ref{eq:rejeita-banda} respectively, considering different cutoff frequencies, center frequencies and band width.}
        \label{fig:iir}
\end{figure}

\begin{equation}\label{eq:passa-baixas}
 \begin{split}
 x & =e^{-2\pi f_c} \\
 a_0 & =  1-x \\
 b_1 & =  x
 \end{split}
\end{equation}

  \item High-pass filter with simple pole and modules of its frequency responses at the upper right corner of Figure~\ref{fig:iir}. The general equation with cutoff frequency $f_c \in (0,\frac{1}{2})$ is calculated by means of the intermediate variable $x \in [e^{-\pi},1]$:

\begin{equation}\label{eq:passa-altas}
 \begin{split}
 x & =e^{-2\pi f_c} \\
 a_0 & =  \frac{x+1}{2} \\
 a_1 & =  -\frac{x+1}{2} \\
 b_1 & =  x
 \end{split}
\end{equation}

%\item Passa-banda
%\item Rejeita-banda

\item Notch filter. This filter is parametrized by a center frequency\footnote{Attention with the cutoff frequency $f_c$ in low-pass and high-pass filters.} $f_c$ and band width $bw$ -- $f_c \pm bw$ that results in $0.707$ of the amplitude, i.e. attenuation of $3dB$ -- both given as fractions of $f_a$, therefore $f,\; bw \in (0,0.5)$.

In order to simplify, consider the auxiliary variables $K$ and $R$ defined as:

\begin{equation}\label{eq:varAux}
 \begin{split}
  R & = 1 - 3bw \\
  K & = \frac{1-2R\cos(2\pi f_c) + R^2}{2 - 2 \cos (2 \pi f_c)}
 \end{split}
\end{equation}

The band-pass filter in the lower left corner of Figure~\ref{fig:iir} has the following coefficients in equation~\ref{eq:diferencas}:

\begin{equation}\label{eq:passa-banda}
 \begin{split}
 a_0 & =  1 - K \\
 a_1 & =  2(K-R)\cos (2\pi f_c) \\
 a_2 & =  R^2-K \\
 b_1 & =  2R \cos (2\pi f_c) \\
 b_2 & =  -R^2
 \end{split}
\end{equation}

The coefficients of band-reject filter are:

\begin{equation}\label{eq:rejeita-banda}
 \begin{split}
 a_0 & =  K \\
 a_1 & =  -2K\cos (2\pi f_c) \\
 a_2 & =  K \\
 b_1 & =  2R \cos (2\pi f_c) \\
 b_2 & =  -R^2
\end{split}
\end{equation}

\noindent with the module of its frequency response represented in the lower left corner of the Figure~\ref{fig:iir}.

%\item Biquad: pela especificação de uma frequência central, da qualidade
%e da intensidade do filtro, este filtro é simples e usual para áudio,
%permitindo ajustes mais finos. Diversas receitas podem ser encontradas
%na literatura, recomendamos especialmente as diferentes especificações
%em ~\ref{musicDSP} e ~\ref{dspguide}.

\end{enumerate}

\subsection{Noise}\label{subsec:ruidos}

Generally, sounds without a defined pitch are called noise~\cite{Lacerda}.
They are important parts of the musical sounds with undefined pitch like the noise present in piano notes, violin, etc. Besides that, the majority of percussion instruments do not have defined pitch and their sounds are generally comprehended as noise~\cite{Roederer}. In electronic music, including electro-acoustic and dance genres, noise has diversified use, and commonly characterizes the music style~\cite{Cook}.

The absence of a defined pitch is due to the absence of a harmonic organization in the sinusoidal components that compose the sound.
In this way, there are many ways to generate noise.
The use of random values to generate the sound sequence $T_i$ is an attractive method but the result is generally not useful because it tends to the white noise~\cite{Cook}.

Another way is to generate noise using the desired spectrum, from which it is possible to perform the inverse Fourier transform.
The spectral distribution should be done with care because, if it uses the same phase or phases with strong correlation, the synthesized sound will include considerably amount of concentrated energy in some periods of its duration.

\begin{figure}[htpq!]
     \centering
         \includegraphics[width=\columnwidth]{figures/ruidos}
     \caption{Colors of noise generated by equations~\ref{eq:branco}, \ref{eq:rosa}, \ref{eq:marrom}, \ref{eq:azul} and \ref{eq:violeta}: resulting spectrum and waveforms.}
         \label{fig:ruidos}
\end{figure}

Some noise with static spectrum are listed below. They are called \emph{colors of noise} since they are associated with colors.
Figure~\ref{fig:ruidos} shows the spectrum profile and the corresponding sound sequence side-by-side. All the noise were generated with a same phase, making it possible to observe the contributions of different regions of the spectrum.

\begin{itemize}

 \item The white note has its name because its energy are distributed equally between all the frequencies. It is possible to realize the white noise with the inverse transform of the following coefficients:

\begin{equation}\label{eq:branco}
 \begin{split}
 c_0 & =0 \quad \text{avoiding bias} \\
 c_i & =e^{j.x}\;,\;\; j^2=-1 \;, \;\; x \; \text{random} \; \in \; [0,2\pi]\;,\;\; i \; \in \; \left[1, \, \frac{\Lambda}{2}-1\right] \\
 c_{\Lambda/2} & = 1 \quad\quad \text{(if even $\Lambda$)}\\ 
 c_i & = c_{\Lambda - i}^*\;,\;\; \text{for}\;  i \; > \;  \frac{\Lambda}{2}
 \end{split}
\end{equation}

The value $c_i$, calculated by the exponential, is only an artifice to obtain unitary module and random phase. Besides that, $c_{\Lambda/2}$ is always real (as discussed in the previous section).

 \item The pink noise has a decrease of $3dB$ for octave. This noise is useful for testing electronic devices, and it has prominent presence in nature~\cite{Roederer}.

\begin{multline}\label{eq:rosa}
 %\begin{split}
 \qquad \qquad \qquad f_{\text{min}}  \approx 15 Hz \\
 f_i  = i \frac{f_a}{\Lambda} \;, \;\; \quad i \;\leq\; \frac{\Lambda}{2},\;\; i\;\in\;\mathbb{N}  \\
 \alpha_i  = \left(10^{-\frac{3}{20}}\right)^{\log _2 \left ( \frac{f_i}{f_{\text{min}}} \right )}  \\
 c_i  =0\;,\;\; \forall \; i \; : f_i<f_{\text{min}} \\
 c_i  =e^{j.x} . \alpha_i\;,\;\; j^2=-1 \;, \;\;\ \\ x \;\; \text{random} \; \in \; [0,2\pi]\;,\;\; \forall \; i \; : f_{\text{min}} \le f_i < f_{\lceil \Lambda/2-1 \rceil}  \\
 c_{\Lambda/2}  = \alpha_{\Lambda/2} \quad \text{(se $\Lambda$ par)} \\ 
 c_i  = c_{\Lambda - i}^*\;,\;\; \text{for}\;  i \; > \;  \Lambda/2 \qquad \qquad
 %\end{split}
\end{multline}

The minimum frequency $f_{\text{min}}$ can be chosen based on the human hearing limit, since no one listens to a pitch with a sound component which frequency is bellow $\approx\; 20Hz$.

Other noises can be made from the pink noise procedure by simply modifying some details, specially the equation that defines $\alpha_i$.

  \item The brown noise received this name after Robert Brown, who described the brownian movement\footnote{Although its origin is disparate given its association with the brown color, this noise become established with this specific name.
  Anyway, this association is satisfactory once the white and pink noises are more strident and associated with intense colors~\cite{Cook,guillaume}.}

What characterizes this noise is the decreasing of $6dB$ by octave. In this way, $\alpha_i$ in the set \ref{eq:rosa} is defined as:

\begin{equation}\label{eq:marrom}
 \alpha_i=(10^{-\frac{6}{20}})^{\log _2 \left( \frac{f_i}{f_{\text{min}}} \right )}
\end{equation}

 \item In the blue noise there is a gain of $3dB$ by octave in a band limited by the minimum frequency $f_{\text{min}}$ and the maximum frequency $f_{\text{max}}$. Therefore, the corresponding equation is also based on the equations set~\ref{eq:rosa}:

\begin{equation}\label{eq:azul}
 \begin{split}
 \alpha_i & = (10^{\frac{3}{20}})^{\log _2 \left ( \frac{f_i}{f_{\text{min}}} \right )} \\
 c_i & =0\;,\;\; \forall \; i \; : f_i<f_{\text{min}} \;\; \text{ou} \;\; f_i>f_{\text{max}} \\
 \end{split}
\end{equation}

 \item The violet noise is similar to the blue noise, but its gain is $6dB$ by octave:

\begin{equation}\label{eq:violeta}
 \alpha_i = (10^{\frac{6}{20}})^{\log _2 \left ( \frac{f_i}{f_{\text{min}}} \right )} \;\;, \quad f_{\text{min}} \approx 15 Hz \\
\end{equation}

 \item The black noise has higher losses than $6dB$ for octave:

\begin{equation}\label{eq:preto}
 \alpha_i=(10^{-\frac{\beta}{20}})^{\log _2 \left( \frac{f_i}{f_{\text{min}}} \right )}\;\;, \quad \beta > 6
\end{equation}

 \item The gray noise is defined as a white noise subject to one of the ISO-audible curves. Those curves are obtained by experiments and imperative to obtain $\alpha_i$. An implementation of ISO 226, that is the last revision of those curves, is in the toolbox \massa.\cite{MASSA}

\end{itemize}

This subsection exposed only noises with static spectrum. There are also characterizations for noises with dynamic spectrum during the time, or noises which are fundamentally transient, like clicks and chirps. The former are easily modeled by an impulse relatively isolated, while chirps are not in fact a noise, but a fast scan of some given frequency band~\cite{Cook}.

The noise from equations~\ref{eq:branco}, \ref{eq:rosa}, \ref{eq:marrom},
\ref{eq:azul} and \ref{eq:violeta} are presented in Figure~\ref{fig:ruidos}. The spectrum were build with the same phase and frequency for each coefficient, making it straightforward to observe the contribution of treble harmonics and bass frequencies.

\subsection{Tremolo and vibrato, AM and FM}\label{subsec:tvaf}

Vibrato is a period variation in pitch (frequency) and tremolo is a period variation in volume (intensity)\footnote{Some musical instruments and contexts use different terms. For example, in piano, the called tremolo is a vibrato and a tremolo is the classification used here. The presented definitions are common in contexts regarding music theory and electronic music. In addition, they are based on a broader literature than the one used for an specific instrument, practice or musical tradition~\cite{Lacerda,Harmonia}.}
For the general case, the vibrato is described as follows:

\begin{equation}\label{vbrGamma}
 \gamma_i'=\left \lfloor i f' \frac{\widetilde{\Lambda}_M}{f_a} \right \rfloor
\end{equation}

\begin{equation}\label{vbrAux}
 t_i'=\widetilde{m}_{\gamma_i' \;\% \widetilde{\Lambda}_M}
\end{equation}

\begin{equation}\label{vbrF}
 f_i=f \left ( \frac{f + \mu }{f} \right )^{t_i'}=f . 2^{t_i'\frac{\nu}{12}}
\end{equation}

\begin{multline}\label{vbrGamma2}
 \Delta_{\gamma_i}=f_i\frac{\widetilde{\Lambda}}{f_a} \quad \Rightarrow \quad \gamma_i = \left \lfloor \sum_{j=0}^{i} f_j \frac{\widetilde{\Lambda}}{f_a} \right \rfloor = \\ = \left \lfloor \sum_{j=0}^{i} \frac{\widetilde{\Lambda}}{f_a}f \left ( \frac{f + \mu }{f} \right )^{t_j'}  \right \rfloor= \left \lfloor \sum_{j=0}^{i} \frac{\widetilde{\Lambda}}{f_a}f . 2^{t_j'\frac{\nu}{12}}  \right \rfloor
\end{multline}

\begin{equation}\label{vbrT}
 T_i^{f, vbr(f',\,\nu)}=\left\{ t_i^{f,vbr(f',\,\nu)} \right\}_0^{\Lambda-1}=\left\{ \widetilde{l}_{\gamma_i \%\; \widetilde{\Lambda} } \right\}_0^{\Lambda-1}
\end{equation}

\begin{figure}[h!]
     \centering
         \includegraphics[width=\columnwidth]{figures/vibrato}
     \caption{Spectrogram of a sound with sinusoidal vibrato of $3Hz$ and one octave depth in a $1000Hz$ sawtooth wave (considering $f_a=44.1kHz$).}
         \label{fig:vibrato}
\end{figure}

For the properly realization of vibrato, it is important to pay attention on the following tables (quais tabelas? seriam termos?) and sequences.
\textcolor{red}{Table $\widetilde{M}_i$ with length $\widetilde{\Lambda}_M$ and the sequence with indices $\gamma_i'$ make the sequence $t_i'$ which is the oscillation pattern in the frequency while table $\widetilde{L}_i$ with length $\widetilde{\Lambda}$ and the sequence with indices $\gamma_i$ make $t_i$ which is the sound itself.
Variables $\mu$ and $\nu$ quantify vibrato intensity: $\mu$ is a direct measure of how many Hertz are involved in the upper limit of the oscillation while $\nu$ is the direct measure of how many semitones (or half steps) are involved in the oscillation ($2\nu$ is the number of semitones between the upper and lower pikes of frequency oscillations of the sound $\{t_i\}$ caused by the vibrato).
It is convenient to use $\nu=\log_{2}\frac{f+\mu}{f} $ in this case because the maximum frequency increase is not equivalent to the maximum diminish, but the semitones variations remains.}

Figure~\ref{fig:vibrato} illustrates the spectrogram of an artificial vibrato for a note with $1000Hz$ (between a \emph{B} and a \emph{C}), and which frequency deviation reaches one octave above and one below. Any waveform can be used to generate a sound and a vibrato oscillation pattern given any oscillation frequency and pitch deviation\footnote{The pitch deviation is called 'vibrato depth' and is generally given as semitones or cents, as convenience}.
Those oscillations with precise forms and arbitrary amplitudes are not possible in traditional music instruments and, in this way, it introduces novelty in the artistic possibilities.

Tremolo is similar: $f'$, $\gamma_i'$ and $t_i'$ remains the same.
The amplitude sequence to be multiplied by the original sequence $t_i$ turns in:

\begin{equation}\label{trA}
 a_i=10^{\frac{V_{dB}}{20}t_i' } = a_{\text{max}}^{t_i'}
\end{equation}

\begin{multline}\label{trT}
 T_i^{tr(f')}=\left \{ t_i^{tr(f')} \right \}_0^{\Lambda-1}=\{ t_i . a_i \}_0^{\Lambda-1}= \\ =\left \{t_i .10^{t_i' \frac{V_{dB}}{20}}    \right \}_0^{\Lambda-1}=\left\{t_i . a_{\text{max}}^{t_i'} \right\}_0^{\Lambda-1}
\end{multline}

\noindent where $V_{dB}$ is the oscillation depth in decibels of tremolo and $a_{\text{max}}=10^{\frac{V_{dB}}{20}}$ is the maximum amplitude gain.
The measurement in decibels is pertinent because the maximum increase in amplitude is not equivalent to the related maximum decrease, while the difference in decibels remains.

Figure~\ref{fig:tremolo} shows the amplitude of sequences $\{a_i\}_0^{\Lambda-1}$ and $\{t_i'\}_0^{\Lambda-1}$ for three oscillations of a tremolo with a sawtooth waveform. The curvature is due to the logarithmic progression of the intensity. The tremolo frequency is $1,5Hz$ because $f_a=44,1kHz \; \Rightarrow \; \text{duration} = \frac{i_{\text{max}}=82000}{f_a}= 2s \; \Rightarrow \; \frac{3\text{oscillations}}{2s}=1,5$ oscillations for second ($Hz$).

The music piece \emph{Vibra e treme} explores those possibilities given by tremolos and vibratos, both used in conjunction and independently, with frequencies $f'$, different depths ($\nu$ and $V_{dB}$), and progressive parameters variations \footnote{Tremolos and vibratos occur many times together in a traditional music instrument and voices.}. Aiming a qualitative appreciation, the piece also develops a comparison between vibratos and tremolos in logarithmic and linear scales. Its source code is in the Appendix~\ref{ap:vibra} and available online as part of the \massa \emph{toolbox}.

\begin{figure}[h!]
     \centering
         \includegraphics[width=\columnwidth]{figures/tremolo}
     \caption{Tremolo with a depth of $V_{dB}=12dB$ with a sawtooth waveform as its oscillatory pattern with $f'=1.5Hz$ in a sine of $f=40Hz$ (considering a sample frequency of $f_a=44,1kHz$).}
         \label{fig:tremolo}
\end{figure}

With progressive increasing of $f'$ for a linear proximity of frequency aiming
to hear the phenomena as pitch ($\approx 20Hz$) it (o que?) generates roughness to both
tremolos and vibratos. Those roughness are largely appreciated both in
traditional classical music and current electronic music, specially in the
\emph{Dubstep} genre.  Roughness is also generated by spectral content that
produce beating~\cite{Porres,porres2009}. The sequence \emph{Bela Rugosi}
explores the threshold with concomitants of tremolos and vibratos at the same
voice, with different intensity and waveforms. The respective code is presented in
Appendix~\ref{ap:bela} and available online in \massa.

Increasing even more the frequency makes those oscillations no longer remains
noticeable. In this case, the oscillations are audible as pitch. Then, $f'$,
$\mu$ and the waveform changes together the frequency of original sound $T_i$
in different ways for both tremolos and vibratos.  They are called AM
(\emph{Amplitude Modulation}) and FM (\emph{Frequency Modulation}) synthesis,
respectively.  These techniques are well known, with applications in
synthesizers like \emph{Yamaha DX7}, and even with applications outside music, like
in telecommunications to data transferring by means of electromagnetic waves
(e.g.\ AM and FM radios).

For musical ends, it is possible to understand FM based on the
case of sines and to decompose the signals into their respective Fourier
components (i.e.\ sinusoidal) to more complex cases. In this way, the FM
synthesis performed with a sinusoidal vibrato with frequency $f'$ and depth
$\mu$ in a sinusoidal sound $T_i$ with frequency $f$ generates bands
centered in $f$ and far from each other with a distance of $f'$:

\begin{equation}\label{eq:fmEsp}
\begin{split}
\{t_i'\} & = \left \{ \cos \left [f . 2 \pi \frac{i}{f_a-1} + \mu . sen \left ( f' . 2 \pi \frac{i}{ f_a -1 } \right ) \right ] \right \} = \\
 & = \left \{ \sum_{k=-\infty}^{+\infty} J_k(\mu) \cos \left [ f . 2 \pi \frac{i}{f_a-1} + k . f' . 2 \pi \frac{i}{f_a-1} \right ]  \right \} = \\
 & = \left \{ \sum_{k=-\infty}^{+\infty} J_k(\mu) \cos \left [ (f+k.f') . 2 \pi \frac{i}{f_a-1} \right ]  \right \}
\end{split}
\end{equation}

\noindent where

\begin{multline}\label{eq:Bessel}
J_k(\mu) = \\ = \frac{2}{\pi} \int_0^{\frac{\pi}{2}}\left [ cos \left (\overline{k}\;\frac{\pi}{2} + \mu . \sin w \right ) . cos \left ( \overline{k}\;\frac{\pi}{2} + k . w \right ) \right ] dw \\  \overline{k} = k \% 2 \;\;,\;\; k \in \mathbb{N}
\end{multline}

\noindent is the Bessel function~\cite{BesselCCRMA,JOSFM} which specifies the
amplitude of each component in FM synthesis.

On those equations, the frequency variation introduced by $\{t_i'\}$ does not
respect the geometric progression that follows the pitch perfection but reflects the
equation~\ref{freqLinear}. Equations~\ref{vbrF} are used for FM and are described
in Appendix~\ref{cap:fmam}, where the spectral content of the FM synthesis is
calculated and obtained with oscillations in the logarithmic scale. In
fact, what is attractive in FM is its simple behavior (only with linear
variations in~\ref{eq:fmEsp}).

For the amplitude modulation (AM):

\newcommand{\OneColEqu}[1]{%
\end{multicols}%
\begin{twocolequfloat}%
\begin{equation}
\{t_i'\}_0^{\Lambda-1} =\{(1+a_i) . t_i\}_0^{\Lambda-1} = \left \{ \left [ 1+M.\sin \left ( f'.2\pi\frac{i}{f_a -1} \right ) \right] .P .\sin \left ( f.2\pi\frac{i}{f_a -1} \right ) \right \}_0^{\Lambda-1} = \\ 
                        =  \left\{P.\sin \left( f.2\pi\frac{i}{f_a -1}  \right ) +  \frac{P.M}{2} \left [ \sin \left( (f-f').2\pi\frac{i}{f_a -1}  \right )  + \sin \left( (f+f').2\pi\frac{i}{f_a -1}  \right ) \right ] \right \}_0^{\Lambda-1}
\end{equation}
\end{twocolequfloat}%
\begin{multicols}{2}%
}

%\begin{widetext}
%\begin{multline*}\label{eq:amEsp}
%\begin{split}
%\{t_i'\}_0^{\Lambda-1} =\{(1+a_i) . t_i\}_0^{\Lambda-1} = \left \{ \left [ 1+M.\sin \left ( f'.2\pi\frac{i}{f_a -1} \right ) \right] .P .\sin \left ( f.2\pi\frac{i}{f_a -1} \right ) \right \}_0^{\Lambda-1} = \\ 
%                        =  \left\{P.\sin \left( f.2\pi\frac{i}{f_a -1}  \right %) +  \frac{P.M}{2} \left [ \sin \left( (f-f').2\pi\frac{i}{f_a -1}  \right )  + \sin \left( (f+f').2\pi\frac{i}{f_a -1}  \right ) \right ] \right \}_0^{\Lambda-1}
%\end{split}
%\end{multline*}
%\end{widetext}

The resulting sound is the original one together with the
reproduction of its spectral content below and above the original frequency,
with a distance of $f'$ from $f$. Again, this is obtained by variations in the
linear scale of the amplitude. Appendix~\ref{cap:fmam} has an exposition of the
spectrum considering an AM performed with an oscillation in the amplitude logarithmic scale. It also loses the simple behavior.

The sequence $T_i$ with frequency $f$, called 'carrier signal', is modulated by
$f'$, called 'modulation signal'. Assuming the FM and AM jargon, $\mu$ and
$\alpha=10^{\frac{V_{dB}}{20}}$ are called 'modulation indexes'. The following
equations are defined for the vibration pattern of the modulation signal
$\{t_i'\}$:

\begin{equation}\label{fmGammaAux}
\gamma_i'=\left \lfloor i f' \frac{\widetilde{\Lambda}_M}{f_a} \right \rfloor
\end{equation}

\begin{equation}\label{fmAux}
t_i'=\widetilde{m}_{\gamma_i' \;\% \widetilde{\Lambda}_M}
\end{equation}

The modulation signal $\{t_i'\}$ into the carrier signal $\{t_i\}$ for FM is applied as:

\begin{equation}\label{fmF}
f_i=f + \mu . t_i'
\end{equation}

\begin{equation}\label{fmGamma}
\Delta_{\gamma_i}=f_i\frac{\widetilde{\Lambda}}{f_a} \quad \Rightarrow \quad \gamma_i = \left \lfloor \sum_{j=0}^{i} f_j \frac{\widetilde{\Lambda}}{f_a} \right \rfloor = \left \lfloor \sum_{j=0}^{i} \frac{\widetilde{\Lambda}}{f_a}(f+\mu . t_j') \right\rfloor
\end{equation}

\begin{equation}\label{fmT}
T_i^{f,\, FM(f',\,\mu)}=\left\{ t_i^{f,\,FM(f',\,\mu)} \right\}_0^{\Lambda-1}=\left\{\,\widetilde{l}_{\gamma_i \%\; \widetilde{\Lambda} } \,\right\}_0^{\Lambda-1}
\end{equation}

\noindent where $\widetilde{l}$ is the waveform period with a length
$\widetilde{\Lambda}$ for the carrier signal.

To perform AM someone just needs to modulate the signal $\{t_i\}$ with $\{t_i'\}$ using
the following equations:

\begin{equation}\label{amA}
a_i=1 + \alpha . t_i'
\end{equation}

\begin{multline}\label{amT}
T_i^{f,\,AM(f',\,\alpha)}=\left\{ t_i^{f,\,AM(f',\,\alpha)} \right\}_0^{\Lambda-1}=\{ t_i . a_i \}_0^{\Lambda-1}= \\ \{t_i . (1 + \alpha . t_i')    \}_0^{\Lambda-1}
\end{multline}

\subsection{Musical usage}\label{subsec:mus2}

At this point, the musical possibilities have been exploded. All characteristics like pitch (given by frequency), timbre (given by the waveform and filters), volume (given by intensity) and duration (given by the number of samples) can be considered by themselves or treated during their duration (with the exception of the duration itself).

The following musical usages comprehend a collection of possibilities
with the purpose of exemplifying types of sound manipulation that results in
musical material. Some of them are deeply discussed in the next section.

\subsubsection{Relations between characteristics}

Another interesting possibility is to use the relations between parameters of
tremolo and vibrato, and some parameters of the basic note like frequency. In
this way, it is possible to say that vibrato frequency is proportionally
different from pitch, or the tremolo depth is inversely proportional to
pitch. Therefore, with equations \ref{vbrGamma}, \ref{vbrF} and \ref{trA}, it is
possible to define:

\begin{equation}\label{eq:vinculos}
\begin{split}
f^{vbr} = f^{tr} & = func_a(f) \\
\nu & = func_b(f) \\
V_{dB} & = func_c(f)
\end{split}
\end{equation}

\noindent with $f^{vbr}$ and $f^{tr}$ as $f'$ in the referenced equations. They can also be associated with the vibrato and tremolo oscillation frequency in
equation~\ref{vbrGamma}.

Besides that, $\nu$ and $V_{dB}$ are the respective depth values of vibrato and
tremolo. Functions $func_a$, $func_b$ and $func_c$ are arbitrary and dependents
on the musical intentions. The music piece \emph{Tremolos, vibratos e a
frequência} explores such characteristics and shows variations in the oscillation
waveform with specific relations with the purpose of building a \emph{musical
language}\footnote{Details in the next section.}. The corresponding code is on
Appendix~\ref{ap:tremolos} and it is also available online as part of the \massa
toolbox.

About convolution it is possible to use a musical pulse as duration -- like a BPM
pulse -- and to distribute impulses during this original pulse, aiming at
establishing metrics and rhythms\footnote{It is important to remember that the
convolution with a impulse results in a sound shifted to the moment that the
impulse occurred}.
For example, two impulses equally spaced builds a binary division into the
pulse. Two signals, one with 2 pulses and another with 3 pulses, both with
equally spaced impulses in the pulse duration, results in the pulse
maintenance with a rhythmic mark that is possible to use as binary or ternary
divisions in many ethnic or traditions music styles~\cite{Gramani}. 
The absolute values of the impulses results in proportions among the
amplitudes of the convoluted signals.
Using convolution with impulses as a metric is explored in the
music piece \emph{Trenzinho de caipiras impulsivos}. The features
embrace the creation of 'sound amalgam' based on granular synthesis and this
piece provides a link to the next section. Refer specially to the
Figure~\ref{fig:pulsoSubAgl}. The source code of the music piece is included in the section~\ref{ap:trenzinho} and online as well (\massa toolbox).

\subsubsection{Moving audio source and receptor, Doppler effect}

Resuming the exposition in subsection~\ref{subsec:spac}: when an audio source
and a receptor are moving, their characteristics are ideally updated at each
sample of the digital signal. The speeds are decomposed in relation of each ear
direction. In this way, given the audio source speed (or velocity) $v_s$, that
is positive if the source moves away from receptor, and receptor speed $v_r$,
that is positive when it gets closer of the audio source, the frequency is given
by the well known Doppler effect:

\begin{equation}\label{eq:fDoppler}
    f=\left(\frac{v_{sound}+v_r}{v_{sound}+v_s}\right)f_0
\end{equation}

With this frequency and relations given by the new IID from the new source
position, it is possible to create the Doppler effect. There is an addendum to
improve the fidelity of the physical phenomena: to increase the received
potency. It is possible to understand this potency gain as being proportional to the relative speed that, in each second, adds the traversed period, by means of a waveform with potency: $\Delta P=P_0\left(\frac{v_r-v_s}{343,2}\right)$, where $P_0$ represents the signal potency.

In this way it is possible to obtain both amplitude and frequency of a moving
audio source. Being this audio source in front of receptor with $y_0$ meters of horizontal distance and $z_0$ meters of height, the distance is given by
$D_i=\left\{ d_i=\sqrt{ y_{i}^{2}+z_{0}^{2} } \right\}_0^{\Lambda-1}$,
where $y_i=y_0+v_s-v_r$ if considering $v_s$ and $v_r$ both in
horizon. Amplitude changes with the distance and with the potency factor cited
above (see subsection~\ref{subsec:volume} for potency to amplitude conversion).

\begin{equation}\label{eq:aDoppler}
    A_i=\left\{ \frac{z_0}{d_i}A_{\Delta P}\right\}_0^{\Lambda-1} = \left\{ \frac{z_0}{\sqrt{y_i^2+z_0^2}} \sqrt{\frac{v_r-v_s}{343,2}+1}  \,\right\}_0^{\Lambda-1}
\end{equation}

Observe that the factor which changes the amplitude (caused by the distance) is even, while the factor caused by the potency variation is antisymmetric in
relation to the crossing of source with receptor. The frequency has a symmetric progression in relation to pitch. In other words, the same semitones (or fractions) added during the approach are decreased during the departure. Besides that, the transition becomes steep if both source and receptor intersects each other exactly at the same point, otherwise, there is a monotonic progression. In the given case, where there is a static height $z_0$, it is necessary to observe the speed component in the direction between the observer and audio source:

\begin{equation}\label{eq:ffDoppler}
    F_i=\{f_i\}_0^{\Lambda-1}=\left\{\frac{v_{sound} + v_r\frac{y_i}{\sqrt{z_0^2+y_i^2}}}{v_{sound}+v_s\frac{y_i}{\sqrt{z_0^2+y_i^2}}}f_0\right\}_0^{\Lambda-1}
\end{equation}

In the Appendix~\ref{sec:cod2} there is a Python implementation of the Doppler
effect as describe above, also considering the intersection between audio source and receptor.

\subsubsection{Filter and noises (subsections~\ref{subsec:ruidos} and~\ref{subsec:filtros})}

Using filters the possibilities are many. It is possible to
convolve a signal to reverberate it, to remove its noise, to generate
distortions or to treat the audio with aesthetically manipulation in mind. For
example, it is possible to simulate sounds from and old television if a
band-pass filter is applied to accept just sound between $1kHz$ and $3kHz$. Or
if someone removes (with some precision) just the frequency of electric
oscillation (usually $50Hz$ or $60Hz$) and the harmonics, it will remove noises
caused by audio devices. A more musical application is to perform filtering in
specific bands and to use those bands as an additional parameter to the notes.

Inspired by traditional music instruments, it is possible to apply a
time-dependent filter~\cite{Roederer}. Chains of those filters can perform
complex and more accurate filtering routines. The music piece \emph{Ruidosa
faixa} explores those features by using filters and many kinds of noise
synthesis. The source code is in Appendix~\ref{ap:ruidosa} and is available
online as part of \massa.

When used together the features can create an effect known
as \emph{chorus}. Based on what happens in a singers choir, in this
effect the sound is performed using small and potentially arbitrary
modifications of parameters like center frequency, presence (or absence) of
vibrato or tremolo and its characteristics, equalization, volume, etc. As a
final result, those versions of the original sound are mixed together (see
equation~\ref{eq:mixagem}). The music piece \emph{Chorus infantil} implements a
chorus in many ways with different sounds and its source code is listed in
Appendix~\ref{ap:chorus}. The script is also available in the \massa toolbox.

\subsubsection{Reverberation}\label{subsubsec:reverb}

Using the same spatialization terms of subsection~\ref{subsec:spac}, the
late reverberation can be modeled as a convolution with a period of pink, brown
or black noise, with exponential decay of amplitude and time related. In
this way, the treble attenuation and irregular smooth in the frequency response
are contemplated with great success. Delay lines can be added as prefix to noise
with the decay, and this contemplates both time parts of the reverberation: the
first reflections and the late reverberation. It is possible to improve the ??
quality by calculating the geometric localization of the last surface where each
wave front reflected before reach the ear in the fairest $100-200$ milliseconds. Then, one could apply a low-pass filter as described in subsection~\ref{subsec:filtros}. The colored
noise can be gradually introduced since the initial moment gives direct
incidence (i.e.\ without any reflection and given by ITD and IID),
with \emph{fade-in}, reaching its maximum at the beginning of the 'late
reverberation', when the geometric incidences loses their importance to the
statistics properties of the noise decay.

As an example, consider $\Delta_1$ as the duration of the first period and $\Delta_R$ as the duration of total reverberation ($\Lambda_1=\Delta_1 f_a$, $\Lambda_R=\Delta_R
f_a$). It is possible to add a probability $p_i$ of a sound reincidencing in the
$i$-th sample with amplitude with exponential decay. Following
subsection~\ref{subsec:spac}, the reverberation $R_i^1$ of the first period can
be described as:

\begin{multline}\label{eq:p1rev}
    R_i^1=\left\{r_i^1\right\}_0^{\Lambda_1-1}\;: \\ \;r_i^1=\left\{
        \begin{array}{l l}
            10^{\frac{V_{dB}}{20}\frac{i}{\Lambda_R-1}}\;  & \text{with probability}\quad p_i=\left(\frac{i}{\Lambda_1}\right)^2 \\
                                     0 \; & \text{with probability}\quad 1-p_i \\
        \end{array} \right.
\end{multline}

\noindent where $V_{dB}$ is the total decay in decibels, typically $-80dB$ or
        $-120dB$. Reverberation $R_i^2$ of the second period can be emulated by a
        brown noise $R_i^m$ (or by a pink noise $R_i^r$) with exponential decay:

\begin{equation}\label{eq:p2rev}
    R_i^2=\left\{r_i^2\right\}_{\Lambda_1}^{\Lambda_R-1}=\left\{10^{\frac{V_{dB}}{20}\frac{i}{\Lambda_R-1}}\,.\,r_i^m\right\}_{\Lambda_1}^{\Lambda_R-1}
\end{equation}

\noindent like:

\begin{equation}\label{eq:rev}
    R_i=\left\{r_i\right\}_0^{\Lambda_R-1}\;:\;r_i=\left\{
        \begin{array}{l l}
            1\; & \text{if }\quad i=0 \\
            r_i^1\;  & \text{if }\quad 1\leq i<\Lambda_1-1 \\
                                     r_i^2 \; & \text{se}\quad \Lambda_1 \leq i < \Lambda_R-1 \\
        \end{array} \right.
\end{equation}

\noindent the reverberation represented by $R_i$ can be applied as simple
        convolution of $R_i$ (called reverberation 'impulse response') with the sound sequence $T_i$ as described in subsection~\ref{subsec:filtros}.

Reverberation is well known by causing great interest in the listeners and to provide more enjoyable sonorities. Besides that, modifications in the reverb
space consists of a tricking (almost a \textit{clich\'{e}}) to induce surprise and interest in the listener.

\subsubsection{ADSR Envelopes}

The volume variation along the duration of sound is crucial to our timbre
perception. The volume envelope, known as ADSR (\emph{Attack-Decay-Sustain-Release}), has many implementations in both hardware and software synthesizers. A pioneer implementation can be found in the Hammond
Novachord synthesizer of 1938 and some variants are cited below~\cite{ADSR}.

The stochastic ADSR envelope is characterized by 4 parameters: attack duration
(time to sound reaches its maximum volume), decay duration (follows the attack
immediately), level of sustain volume (in which the volume remains stable after the decay) and release duration (after sustain, this is the duration until the volume decays to zero). Note that the sustain duration is not specified because it is the difference between the duration itself and the durations of attack, decay and sustain.

It is possible to apply the ADSR envelope with durations $\Delta_A$, $\Delta_D$ and $\Delta_R$, with total duration $\Delta$ and sustain level $a_S$, given by the fraction of the maximum amplitude, as a sound sequence $t_i$ defined as:

\begin{multline}\label{eq:adsr}
%\begin{split}
\{a_i\}_0^{\Lambda_A-1}  = \left\{\xi\left(\frac{1}{\xi}\right)^{\frac{i}{\Lambda_A-1}}\right\}_0^{\Lambda_A-1} \;\; \text{or} \;\;\quad \left\{\frac{i}{\Lambda_A-1}\right\}_0^{\Lambda_A}\\
\{a_i\}_{\Lambda_A}^{\Lambda_A+\Lambda_D-1} =\left\{a_S^{\frac{i-\Lambda_A}{\Lambda_D-1}}  \right\}_{\Lambda_A}^{\Lambda_A+\Lambda_D-1} \;\;\quad \quad\quad \\ \text{or} \quad\;\; \left\{1-(1-a_S)\frac{i-\Lambda_A}{\Lambda_D-1}\right\}_{\Lambda_A}^{\Lambda_A+\Lambda_D-1}\\
\{ a_i \}_{\Lambda_A+\Lambda_D}^{\Lambda-\Lambda_R-1} =\left\{ a_S \right\}_{\Lambda_A+\Lambda_D}^{\Lambda-\Lambda_R-1} \\
\{ a_i \}_{\Lambda-\Lambda_R}^{\Lambda-1}  =\left\{ a_S\left(\frac{\xi}{a_S} \right)^{\frac{i-(\Lambda-\Lambda_R)}{\Lambda_R-1}} \right\}_{\Lambda-\Lambda_R}^{\Lambda-1} \quad\;\; \\ \text{or} \quad\;\; \left\{ a_S - a_S\frac{i+\Lambda_R-\Lambda}{\Lambda_R-1}\right\}_{\Lambda-\Lambda_R}^{\Lambda-1} 
%\end{split}
\end{multline}

\noindent with $\Lambda_X=\lfloor \Delta . f_a \rfloor\;\;\forall\;\; X \; \in
(A,D,R,\;)$ and being $\xi$ a small value that provides a satisfactory \emph{fade in} and \emph{fade out}, e.g.\ $\xi=10^{\frac{-80}{20}}=10^{-4}\;$ or $\;\xi=10^{\frac{-40}{20}}=10^{-2}$. The lower the $\xi$, the slower the \emph{fade}, like the $\alpha$ illustrated in Figure~\ref{fig:transicao}. The terms in the right side of the equation~\ref{eq:adsr} attend for both introduction and ending of sound from the zero intensity because they are linear. Schematically, Figure~\ref{fig:adsr} shows the ADSR envelope, a classical implementation that supports many variations. For example, between attack and decay it is possible to add an extra partition where the maximum amplitude stays. Another common example is the use of more elaborated tracings to attack or decay. The music piece \emph{ADa e SaRa}, available both in
Appendix~\ref{ap:ada} and in \massa, explores many configurations of the ADSL envelope.

\begin{equation}\label{eq:adsrApl}
\left\{t_i^{ADSR}\right\}_0^{\Lambda-1} =\{t_i . a_i\}_0^{\Lambda-1}
\end{equation}

\begin{figure}[htpq!]
    \centering
        \includegraphics[width=\columnwidth]{figures/adsr}
    \caption{An ADSR envelope (\emph{Attack, Decay, Sustain, Release}) applied
        to an arbitrary sound sequence. The linear variation of the amplitude is
        above. Below the amplitude variation is exponential.}
        \label{fig:adsr}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\afterpage{\blankpage}

%\clearpage

\section{Notes organization in music}\label{notasMusica} \label{sec:notesMusic}

Consider $ S_j=\left\{  s_j=T_i^j=\{t_i^j\}_{i=0}^{\Lambda_j-1} \right\}_{j=0}^{H-1} $ as the sequence $S_j$ of $H$ musical events $s_j$. In addition, $S_j$, called a 'musical structure', as composed by events $s_j$ which are musical structure themselves, e.g.\ notes. This section is dedicated to techniques that make $S_j$ interesting and enjoyable for audition.

The elements of $S_j$ can be overlapped by mixing them together as in
equation~\ref{eq:mixagem} and Figure~\ref{fig:mixagem}, building intervals
and chords. This scenery reflects the 'vertical thought' in music. On the other hand, the concatenation of $S_j$
as in equation~\ref{eq:concatenacao} and in Figure ~\ref{fig:concatenacao},
builds melodic sequences and rhythms, which are associated with the musical
'horizontal thought' in music. The fundamental frequency $f$ and the beginning moment
(attack) are generally the most important characteristics of the elements in
$S_j$. Such characteristics make possible the creation of music made by pitches (both
harmony and melody) and by the presence of temporal metrics and rhythms.

\subsection{Tuning, intervals, scales and chords}\label{subsec:afinacao}

Here we describe characteristics in music related to harmony and melody.

\subsubsection{Tuning}

Doubling the frequency is equivalent to ascending one octave ($f=2f_0$).
The octave division in twelve notes is the canon of classical western music, although its usage has also been observed
outside western tradition, like ceremonial/religious and ethnic context~\cite{Wisnick}.

The factor given by $\varepsilon=2^{\frac{1}{12}}$ defines a semitone. It builds a note
grid along the hearing spectrum. Fixing a frequency $f$, all the possible
fundamental frequencies stay separated by intervals that are multiples of
$\varepsilon$. Twelve semitones (or half steps), equidistant to the human ear, make an
octave. Therefore, if $f=2^{\frac{1}{12}}f_0$, there is a semitone between
$f_0$ and $f$.

This absolute accuracy is common in computational implementations. For the real
music instruments, however, there are deviations of those frequencies in order to make
the harmonics of corresponding notes more compatible. Besides that, the fixed reference
$\varepsilon=2^{\frac{1}{12}}$ characterizes an equally tempered tuning. There
are tunings which respective intervals are ratios of low-order integers, originally based
on observations of physical behaviors. First of these tunings were formalized
around two thousand years before the existence of an equal temperament~\cite{Roederer}.

Two iconic tunings:

\begin{itemize}
    \item The {\bf just intonation} is defined as notes of the diatonic scale
    that are associated with ratios of low-order integers that defines the harmonic
    series. The basic ratios are the same in the Ionian mode (white piano keys
    from \do to \do (see below in subsection~\ref{subsec:escalas}): 1, 9/8, 5/4,
    4/3, 3/2, 5/3, 15/8, 2/1). 
    The intervals are considered as notes in a scale together with the following: the
    semitone 16/15, the 'minor tone' 10/9, and the 'major tone' 9/8. There are
    many ways to perform the division of the 12 notes.

    \item The {\bf Pythagorean tuning} is based on the interval 3/2 (perfect
    fifth). The Ionian mode becomes: 1, 9/8, 81/64, 4/3, 3/2, 27/16, 243/128,
    2/1. The intervals are also considered as notes in a scale. Besides the
    intervals in the mode, it is also used the minor second 256/243, the minor
    third 32/27, the augmented fourth 729/512, the diminished fifth 1024/729,
    the minor sixth 128/81 and the minor seventh 16/9. 
\end{itemize}

In order to attends for micro-tonality\footnote{Micro-tonality is the use of intervals less
than one semitone and have ornamental and structuring functionalities in music. The
division of the octave in $12$ notes is motivated by physical fundamentals but it is also
a \emph{convention} adopted even by western classical music. Other tunings are
present. To cite one example, the traditional Thai music uses a octave
division in seven notes equally spaced ($\varepsilon=2^{\frac{1}{7}}$),
resulting in intervals that have little in common with the intervals in the
diatonic scale~\cite{Wisnick}}. It is possible to use non-integer real values as
pitch sequences or even to maintain the integer values and change the factor $\varepsilon=2^{\frac{1}{12}}$. For example, 
a tuning really near of the harmonic series
itself is proposed with the division of the octave in $53$ notes:
$\varepsilon_2=2^{\frac{1}{53}}$~\cite{microtonalidade}.
In that division the notes are related by means of integers with
$\varepsilon_2$. Note that if $S_i$ is a pitch sequence related by means of
$\varepsilon_1$, a mapping for all the notes related by $\varepsilon_2$ defines
a new sequence $S_i'=\left\{s_i'\right\}=\left\{
s_i \frac{\varepsilon_1}{\varepsilon_2}\right\}$. The music piece \emph{Micro
tom} explores this micro-tonal features and its code is in
Appendix~\ref{ap:micro} and is part of \massa.

\subsubsection{Intervals}\label{subsec:intervalos}

Using the ratio $\varepsilon=2^{\frac{1}{12}}$ between the note frequencies
(i.e.\ one semitone) the intervals in the twelve note system can be represented by integers. Table~\ref{eq:intervalos} summarizes the characteristics of each interval: its traditional nomenclature, consonance and
dissonance properties, and number of semitones per interval.

\begin{table*}[htpq!]
\centering
\caption{The music intervals together with their traditional notation, basic classification for dissonance and consonance, and number of semitones. Perfect (P) unison, fifths and octaves are the perfect consonances. Major (M) and minor (m) thirds and sixths are the imperfect consonances. Minor seconds and major sevenths are the harsh dissonances. Major seconds and minor sevenths are the mild dissonances. Perfect fourth is the first special case, being a perfect consonance when considering it as the inversion of the perfect fifth and a dissonance or a imperfect consonance otherwise. The second special case is the tritone (A4 or aug4, d5 or dim5, tri,TT). This interval is consonant in some cultures. For tonal music, the tritone indicates dominant and searches for its resolution in a third or sixth. Due to this instability it is considered a dissonance interval.}
\begin{tabular}{| c | c | c | }\hline
    \multicolumn{3}{|c|}{\bf consonances}  \\\hline
   & traditional notation & number of semitones \\
   perfects: & P1, P5, P8 & 0, 7, 12 \\
 imperfects: & m3, M3, m6, M6 & 3, 4, 8, 9 \\\hline\hline
    \multicolumn{3}{|c|}{\bf dissonances} \\\hline
 & traditional notation & number of semitones \\
 fortes: & m2, M7 & 1, 11 \\
 brandas: & M2, m7 & 2, 10 \\\hline\hline
    \multicolumn{3}{|c|}{\bf special cases} \\\hline
 & traditional notation & number of semitones \\
 consonance or dissonance: & P4 & P5 \\
 dissonance in Western tradition: & tritone, aug4, dim5 & 6 \\\hline
\end{tabular}\label{eq:intervalos}
\end{table*}

The nomenclature is based both in impositions and for convenience of the tonal system and practical aspects of note manipulation and can be specified
as~\cite{Roederer,Wisnick}:

\begin{itemize}
                \item Intervals as number of ratios between notes: first
                (unison), second, third, forth, fifth, sixth, seventh and
                eighth. The ninth, tenth, eleventh, etc, are compound                       intervals made by one or more octaves + a interval inside the                 octave, which characterizes the compound interval. The                        intervals are represented by the numeric digits, e.g. 1, 3, 5                 are a unison, a third and a fifth, respectively.

                \item Qualities of each interval: the perfect consonances --
                i.e.\ unison, forth, fifth and octave -- are 'perfect'. The
                imperfect consonances -- i.e.\ thirds and sixths -- and
                dissonances -- i.e.\ seconds and sevenths -- can be major and
                minor. Exception for the tritone.

                \item The perfect fourth is both perfect consonance or
                dissonance according to the context and theoretical
                background. As general rule, it can be considered as                        consonance except when it is used as passage to resolve a                   fifth or a third.

                \item Tritone is dissonance in Western music because
                it characterizes the ``dominant'' in the tonal system (see
                subsection~\ref{subsec:harmonia}) and represents the
                instability. Some cultures chant the interval as consonance.

                \item A major interval decreased by one semitone results in a
                minor interval. A minor interval increased by one semitone
                results in a major interval.

                \item A perfect interval (unison, perfect forth, perfect                    fifth, perfect octave), or a major interval (major second M2,                 major third M3, major sixth M6 or major seventh M7), increased                 by one semitone results in an augmented interval (e.g.\                    augmented third aug3 with five semitones). The augmented forth                 is also called tritone (aug4 ~ tri ~ TT).

                \item A perfect interval or a minor interval (minor second m2,
                minor third m3, minor sixth m6 or minor seventh m7), decreased
                by one semitone results in a diminished interval. The                       diminished fifth is also called tritone (dim5 ~ tri ~ TT).

                \item An augmented interval increased by one semitone results                 in a 'doubly augmented' interval and a diminished interval
                decreased by one semitone results in a 'doubly diminished'
                interval.

                \item Notes played simultaneously configure a harmonic
                interval.

                \item Notes played as a sequence in time configure a
                melodic interval. The order of the notes: considering first                 the lowest note or the highest note, results in an ascending                 or descending interval, respectively.

                \item If the lower pitch is raised one octave, or if
                the upper pitch is lowered one octave below, the interval is
                inverted. When summed to its inversion an interval results in                 9 (e.g.\ m7 is inverted to M2: $m7+M2=9$). An inverted major
                interval results in a minor interval and vice-versa. An                     inverted augmented interval results in a diminished interval                 and vice-versa (like a doubly augmented results in a doubly
                diminished and vice-verse). An inverted perfect interval                    results in a perfect interval as well.

                \item An interval higher than an octave is called a 'compound
                interval' and is classified like the interval between the same
                notes but in the same octave. They are also specified adding a                 7 to the interval: P11 is an octave plus a forth ($7 + P4 =                  P11$), M9 is an octave plus a major second ($7 + M2 = M9$).
\end{itemize}

The augmented/diminished intervals and the doubly augmented/doubly diminished intervals are consequences of the tonal system. The scale degrees (subsection~\ref{subsec:escalas}) generally determine different pitches, with
specific uses and functions. Henceforth, in a \textit{C flat} major scale, the tonic -- first degree -- is \textit{C flat}, not \textit{B}, and the leading tone -- seventh degree -- is \textit{B flat}, not \textit{A sharp} or \textit{C double flat}. In a similar way, the second degree of a scale can be one semitone far from the first degree like the leading tone (seventh degree at one ascending semitone from the first degree), where there is a diminished third between the two semitones of the seventh and second scale degrees, consequence of the first degree be between the near two degrees: the second and the leading note~\cite{Lacerda}.

The descriptions presented here summarizes the traditional theory of the music
intervals~\cite{Lacerda}. The music piece \emph{Intervalos entre alturas}
explores these intervals in a independently and conjunction manner. The reader is invited to see the source code in Appendix~\ref{ap:intervalos} and available online in the \massa toolbox~\cite{MASSA}.

\subsubsection{Scales}\label{subsec:escalas}

Basically speaking, scale is a set of ordered pitches. Usually, scales repeat at each octave. The ascending sequence with all notes from the octave division in 12 equal intervals, separated by ratio $\varepsilon=2^{\frac{1}{12}}$, is known as the chromatic scale with equal temperament. There are 5 perfect symmetric divisions of the octave within the chromatic scale. These divisions are considered as scales by the easily and peculiar use they provide. Considering the integers that the factor $\varepsilon=2^{\frac{1}{12}}$ is powered to multiply $f_0$, the scales are the following:

\begin{multline}\label{escSim}
%\begin{split}
\text{chromatic}  = E_i^c = \\ = \{e_i^c\}_0^{11} =  \{0,1,2,3,4,5,6,7,8,9,10,11\} = \{i\}_0^{11}\\
\text{whole tones}  = E_i^t = \{e_i^t\}_0^{5} = \{0,2,4,6,8,10\} = \{2.i\}_0^{5} \\
\text{minor thirds}  = E_i^{tm} = \{e_i^{tm}\}_0^{3} = \{0,3,6,9\} = \{3.i\}_0^3 \\
\text{major thirds}  = E_i^{tM} = \{e_i^{tM}\}_0^{2} = \{0,4,8\} = \{4.i\}_0^2\\
\text{tritones}  = E_i^{tt} = \{e_i^{tt}\}_0^{1} = \{ 0, 6 \} = \{6.i\}_0^1
%\end{split}
\end{multline}

Therefore, the third note of the whole tone scale with $f_0=200Hz$
is $f_3=\varepsilon^{e_3^t} . f_0 = 2^{\frac{6}{12}} . 200 \approxeq 282.843
Hz$. These 'scales' or patterns generate stable structures by their intern
symmetries and can be repeated in a efficient and sustained way. Section~\ref{estCic} specially discusses the symmetries. 
The music piece \emph{Cristais} exposes each one of these scales, in both melodic and
harmonic counterpart and the corresponding source code is presented in Appendix~\ref{ap:cristais} and as
part of \massa.

The diatonic scales are:

\begin{multline}\label{eq:escalas}
%\begin{split}
\text{natural minor} = \text{aeolian mode}  = \\ = E_i^m = \{e_i^m\}_0^6 = \{0,2,3,5,7,8,10\} \\
\text{locrian mode}  = E_i^{mlo} = \{e_i^{mlo}\}_0^6 = \{0,1,3,5,6,8,10\} \\ 
\text{major}  = \text{ionian mode}  =  E_i^M = \{e_i^M\}_0^6 = \{0,2,4,5,7,9,11\} \\
\text{dorian mode}  = E_i^{md} = \{e_i^{md}\}_0^6 = \{0,2,3,5,7,9,10\} \\
\text{phrygian mode}  = E_i^{mf} = \{e_i^{mf}\}_0^6 = \{0,1,3,5,7,8,10\} \\
\text{lydian mode}  = E_i^{ml}=\{e_i^{ml}\}_0^6 = \{0,2,4,6,7,9,11\} \\
\text{mixolydian mode} = E_i^{mmi} = \{e_i^{mmi}\}_0^6 = \{0,2,4,5,7,9,10\}
%\end{split}
\end{multline}

\noindent They have only the major, minor and perfect intervals. The unique exception is the tritone that is presented as a augmented forth or diminished fifth.

All the diatonic scales follow a pattern of successive intervals \textit{tone,
tone, semitone, tone, tone, tone, semitone}. Thus, it is possible to write:

\begin{equation}\label{eq:relacaoDia}
\begin{split}
\{d_i\} & =\{2,2,1,2,2,2,1\} \\
e_0 & =0 \\
e_i & =d_{(i+\kappa)\%7}+e_{i-1} \quad for \;\;  i > 0
\end{split}
\end{equation}

\noindent with $\kappa \in \mathbb{N}$. For each mode there is only one value
for $\kappa \in [0,6]$. For example, a brief inspection reveals that
$e_i^{ml}=d_{(i+2)\%7}+e_{i-1}^{ml}$. Then, $\kappa=2$ for the lydian mode.

The minor scales has two additional forms, named melodic and harmonic:

\begin{multline}\label{eq:escalasMenores}
%\begin{split}
\text{natural minor (same as above)}  = \\ = E_i^m = \{e_i^m\}_0^6 = \{0,2,3,5,7,8,10\} \\
\text{harmonic minor}  = E_i^{mh} = \{e_i^{mh}\}_0^6 = \{0,2,3,5,7,8,11\} \\
\text{melodic minor}  = \\ = E_i^{mm} = \{e_i^{mm}\}_0^{14} = \{0,2,3,5,7,9,11,12,10,8,7,5,3,2,0\} \\
%\end{split}
\end{multline}

The ascending and descending contour of the melodic minor scale is necessary for the existence of the leading tone (seventh and last degree, separated by one semitone of the octave, enhances the tonic polarization) in the ascending trajectory, which is not necessary in the descending mode, since it recovers the normal form. The harmonic scale presents the leading tone but does not avoid the augmented second interval between the sixth and seventh degrees. In addition, it does not need to cover a melodic trajectory, just having to present the leading tone, crucial to the tonal system (the leading tone tends to the tonic, asserting it)~\cite{Harmonia}. 
Other scales can be represented in a similar form, like the pentatonic and the
modes of limited transposition by Messiaen~\cite{Messiaen}.

\subsubsection{Chords}\label{subsec:acordes}

The simultaneous occurrence of three or more notes is observed by means of chords. Their
base of tonal music are the triads. Triads are built by two successive thirds
within 3 notes: root, third and fifth. If the lower note of a chord is different from the root, this chord is referred
as inverted. A close position is where any note of chord fits between two consecutive
notes~\cite{Lacerda}. All the non-inverted triads in the close position form and with root in
$0$ are described as following:

\begin{equation}\label{triades}
\begin{split}
\text{major triad} = A_i^M= \{a_i^M\}_0^2=\{0,4,7\} \\ 
\text{minor triad} = A_i^m = \{a_i^m\}_0^2=\{0,3,7\} \\
\text{diminished triad} = A_i^d = \{a_i^d\}_0^2=\{0,3,6\} \\
\text{augmented triad} = A_i^a = \{a_i^a\}_0^2=\{0,4,8\}
\end{split}
\end{equation}

To consider another third superimposed to the fifth, is sufficient to add $10$
as the highest note in order to form a tetrad with minor seventh, or add $11$ in order to form a tetrad with major
seventh. The inversions and open positions can be obtained with a selective
addition of $12$ to the components.

Incomplete triadic chords, with extra notes ('dirty' chords), and non-triadic
are also common.

In the following we present general recommendations:

\begin{itemize}
        \item A fifth is a root (fundamental) confirmed by the interval.
        \item The major or minor third points to the major or minor quality
        of the chord.
        \item Every tritone, specially if built between a major third and a
        minor seventh, tends to resolve in a third or sixth.
        \item Note duplication is avoided. If duplication is needed, the
        preferred order is: the root, fifth, third and seventh.
        \item It is possible to build chords with notes different from triads,
        specially if they respect to a recurrent logic or musical chain that
        justifies these different notes.
        \item Chords built by successive intervals different of thirds -- like
        fourths and seconds -- are recurrent in compositions with advanced
        tonalism or experimental character.
        \item The repetition of chords successions (or their characteristics)
        fixes a trajectory by means of the recurrence and make possible the
        introduction of exotic formations without incoherence.
\end{itemize}

\subsection{Atonal and tonal harmonies, harmonic expansion and modulation}\label{subsec:harmonia}

Omission of the basic chaining is the key to obtain modal and atonal harmonies. In
case of absence of these minimal tonal structures, we say that the
harmony is modal if the notes match with some diatonic scale (see
equations~\ref{eq:escalas}) or if the notes are presented in a small number.
If basic tonal chaining are absent, the notes do not match any diatonic scale
and are diverse and dissonant (by relation with each other) enough to avoid
reduction by polarization, the harmony is atonal. In this classification, the
modal harmony are not tonal neither atonal. The modal harmony is reduced to the 
incidence of notes within the diatonic scale (or simplifications) and to the
absence of tonal structures. It is possible to note, following this concept, that
atonal harmony is hard to be realized~\cite{harmEXT}.

\subsubsection{Atonal harmony}

In fact, the techniques around atonal music comprehend structures for avoiding
the hearing relation with modes and tonality. This difficulty imposes the use of dodecafonism. 
The purpose of dedocafonism is to use a set of notes (ideally 12 notes), and to perform each note, one by one, in the same
order (não sei se ficou mt claro). In this context, the tonic becomes difficult to settle. Nevertheless, the
Western hearing searches for tonal traces in musics and easily finds them by
unexpected paths and even tortuous.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The use of dissonant intervals (specially tritones without resolution), together
with ninths, seconds and sevenths, reinforces the absence of tonality. Considering
this context, it is possible to account for the following considerations while creating a music
piece:

\begin{itemize}
     \item Repeating notes. Considering the immediate repetition as an extension
     of the previous incidence, the use of same notes in sequence does not add relevant information.
     \item It is interesting to play adjacent notes at the same time, making harmonic intervals
     and chords.
     \item Present durations with liberty, respecting the notes order.
     \item In order to gain variation, one can use extension, transposition, translation, inversion, retrograde and retrograde inversion. See subsections~\ref{subsec:motivos}
     and~\ref{subsec:usosmusicais3} for more details.
     \item Account for the presence of note structures, it is possible to use variations in orchestration, articulation, spatialization, among others.
\end{itemize}

The atonal harmony can be observed, paradigmatically, within these presented
conditions. Most of them were written by great dodecafonic composers like Alban Berg and even Schoenberg. Many pieces create purposeful links between tonal and atonal techniques.

\subsubsection{Tonal harmony}

In the XX century, rhythmic music with emphasis on sonorities/timbres, extended
the concepts about tonality and harmony. Nevertheless, the
tonal harmony has strong influence in the artistic movements and commercial
venues. In addition, the dodecafonism is considered having tonal nature because it
was created to deny the tonal characteristics of polarization.

In tonal or modal music, the base of the harmonic filed are chords -- like the ones listed in
equations~\ref{triades}-- built with a root note for each degree of a scale -- like
the ones in equation~\ref{eq:escalas}. In this context, music harmony aims at studying the observation of incident chord progressions and chaining rules. Even a monophonic melody generates harmonic fields, making it possible to observe the suggested chords at each passage.

In the 'traditional tonal harmony', a scale has its corresponding tonic (first scale
degree) and can be major (with the same notes of the Ionian mode) or
minor (notes from Eolian mode form the 'natural minor', which has both a
harmonic and a melodic version, see equation~\ref{eq:escalasMenores}). The
scale is base on triads, each one with its root and representing a degree into
the scale: $\hat{1},\hat{2},\hat{3},\hat{4},\hat{5},\hat{6},\hat{7}$. 
To build triads, the third and the fifth note above the root is considered together with the root (or fundamental). It is possible to write $\hat{1},\hat{3},\hat{5}$ as the first degree chord, built by the first degree of scale and the center for a tonal music. The chords
of the fifth degree $\hat{5},\hat{7},\hat{2}$ ($\hat{7}$ sharp when a minor
scale) and of the forth degree $\hat{4},\hat{6},\hat{1}$ are secondary. After
that, other degrees are considered. The traditional harmony comprises conventions and stylistic techniques to create chord chains, made in each scale degree~\cite{Harmonia}.

The 'functional harmony' ascribes functions to these three central chords and
tries to understand their use by means of these functions. The chord built under
the first degree is named the \textbf{tonic} chord (\textit{T} or \textit{t} for a major or minor tonic, respectively) and its (role) function consisting of preserving a center, usually referred as a ``ground'' for the music. The chord built under the fifth degree is the \textbf{dominant} (\textit{D}, the dominant is always major) and its function is call for the tonic (we say that the dominant chord asks for a conclusion and this conclusion is usually the tonic chord). Thus, the dominant chord guides the music to the tonic. The triad built under the forth degree is the \textbf{subdominant} (\textit{S} or \textit{s} for a major or minor
subdominant, respectively) and its basically function is to distance the music from the tonic. The process aims at confirming the tonic using tonic-subdominant-tonic chains which are
expanded by using other chords in different ways.

The remaining triads are associated to these three most important chords. In the major scale, the associated relative (relative tonic \textit{Tr}, relative
subdominant \textit{Sr} and relative dominant \textit{Dr}) is the triad built with a
third below, and the associated counter-relative (counter-relative
tonic \textit{Ta}, counter-relative subdominant \textit{Sa} and the
counter-relative dominant \textit{Da}) is the triad built with a third above. In the
minor scale the same happens, but the triad with a third below is called
counter-relative (tA, sA) and the triad with one third above is called relative (tR,
sR). The precise functions and musical effects of these chords are
controversial. Table~\ref{tab:harmonia} shows the relation between the triads
built in each degree of the major scale.

\begin{table}[htpq!]
\centering
\caption{Summary of tonal harmonic functions of the major scale. Tonic is the
music center, the dominant goes to the tonic and the subdominant moves the music away from
the tonic. The three chords can, at first, be freely replaced by their
respective relative and counter-relatives.}
\begin{tabular}{l | c | r}
relative & main chord of the function & counter-relative \\\hline\hline
$\hat{6},\hat{1},\hat{3}$ & tonic:       $\hat{1},\hat{3},\hat{5}$ & $\hat{3}, \hat{5},      \hat{7}$ \\
$\hat{3},\hat{5},\hat{7}$ & dominant:    $\hat{5},\hat{7},\hat{2}$ & [ $\hat{7},\hat{2},\hat{4}\#$ ] \\
$\hat{2},\hat{4},\hat{6}$ & subdominant: $\hat{4},\hat{6},\hat{1}$ & $\hat{6},\hat{1},       \hat{3}$
\end{tabular}
\label{tab:harmonia}
\end{table}

The dominant counter-relative should form a minor chord. It explains the change
in the forth degree by a semitone above $\hat{7}\#$. The diminished chord
$\hat{7},\hat{2},\hat{4}$, is generally considered a 'dominant seventh chord
with no root'~\cite{Koellheuteur}.
In the minor mode, there is the change in $\hat{7}$ by an ascending semitone to
make possible the existence of an unique semitone separating the tonic, and
making also possible the dominant (that should be major and goes to the tonic). In
this way, the dominant is always major, for both major and minor scale and,
because of that, even as a minor tone, the relative dominant remains a third
below and in the counter-relative, a third above.

\subsubsection{Tonal expansion: individual functions and chromatic mediants}

%% dominante e subdominante individualmente, ou, dominante e subdominante individual?

Each one of these chords can be confirmed and developed by performing their
dominant or individual subdominant, which is the chord based in the third formed
by a fifth above or by a fifth below, respectively. These dominants and individual
subdominants, in the same way, have also subdominants and individual dominants that
can be used. Given a tonality, any chord can occur, no matter
how distant it is from the harmonic field and from the notes within the
scale. The unique on condition is that the occurrence presents a coherent trajectory of
dominants and subdominants to the original tonality.

Mediants, or 'chromatic mediants', are present in each chord in two ways: the upper chromatic
mediant, formed by the root at the third of original chord; and the lower
chromatic mediant, formed by the fifth at the third of original chord. Both chords also are formed by a third, but with a chromatic alteration regarding
the original chord. If two chromatic alterations exist, i.e.\ two notes altered
by one semitone each regarding the original chord, the mediant is called
'doubly chromatic mediant'. Again, there are two forms for each chord: the upper form, with
a third in the fifth of the original chord; and the lower form, with a third in the root
of the original triad. It is interesting to observe that a major chord has upper chromatic mediants and upper doubly chromatic mediants. A minor chord has lower chromatic mediants and lower doubly chromatic mediants. This relation between chords is considered as
advanced tonalism, sometimes even considered as expansion and dissolution of tonalism, and, as consequence, it has strong and outstanding effects although perfectly consonant. The chromatic
mediants are used since the end of Romanticism by Wagner, Lizt, Richard Strauss,
among others, and are quite simple to be realized~\cite{Harmonia,Salzer}.

\subsubsection{Modulation}

Modulation is the change of the key (tonic, or tonal center) in a music. It can be characterized by observing the key at the start and at end of the chord transitions. Keys are always conceived as related by fifths and their relative and counter-relatives. Forms to perform modulation include:

\begin{itemize}
    \item Transposing the discourse to a new key, without any preparation. It is
    a common Baroque procedure although incident at other periods as well. Sometimes it is 
    called phrasal modulation or unprepared modulation.
    \item  Using carefully an individual dominant and perhaps the individual
    subdominant, to confirm the change in key and harmonic field.
    \item Using chromatic alterations to reach a chord in the new key starting from a
    chord in the previous key. It is called chromatic modulation.
    \item Featuring a unique note, possibly repeated or suspended with no
    accompaniment, common to start and end a key, constitute a peculiar way
    to introduce the new harmonic field.
    \item Changing the function, without changing the notes, of a chord to
    contemplate a new key. This procedure is called enharmony.
    \item Maintaining the tonal center and changing the key quality from major to minor
    (or vice-verse) comprehends a parallel modulation. Key with same tonic and
    different quality is known as homonyms.
\end{itemize}

The dominant has great importance and is a natural pivot into modulations,
which leads to the circle of fifths~\cite{Harmonia,Salzer,Koellheuteur,Harmony}. 
The music piece \emph{Acorde cedo} explores these chord relations. Its code is
available both in Appendix~\ref{ap:acorde} and online as part of \massa~\cite{MASSA}.

\subsection{Counterpoint}\label{subsec:contraponto}

Counterpoint is defined as the conduction of simultaneous melodic lines, called voices. The bibliography covers systematic ways to conduct voices, leading to scholastic genres like canons, inventions and fugues. It is possible to
summarize the main counterpoint rules, and it is known that Beethoven --
among others -- outlined such similar synthesis.

\begin{figure}[h!]
    \centering
        \includegraphics[width=\columnwidth]{figures/movContraponto}
    \caption{Different motions of counterpoint aiming to preserve independence
        between voices. There are 3 types of motions: direct, contrary and
        oblique. The parallel motion is a type of direct motion.}
        \label{fig:movContraponto}
\end{figure}

The purpose of counterpoint is to conduct voices in a way they sound
independent. In order to do that, the relative motion of voices (in pairs), is crucial and
categorized as: direct, oblique and contrary motion (see Figure~\ref{fig:movContraponto}). The parallel motion is an oblique motion.
The gold rule here is to take care with the direct motions, avoiding making them
lead to a perfect consonance. The parallel motion should occur only between
imperfect consonances and no more than three consecutive times. The dissonances
can be unadmitted or used when followed and preceded by consonances of joint
degrees, i.e.\ neighbor notes in a scale. The motions that lead a note to a
neighbor sound more coherent. When having 3 or more voices, the melodic
importance lies on the higher and lower voices, in this particular
order~\cite{Fux,Tragtenberg,SchoenbergContra}.

These rules were used in the music piece \emph{Conta ponto}, the source code is
in Appendix~\ref{ap:conta} and available online in \massa.

\subsection{Rhythm}\label{subsec:ritmo}

Basically speaking, the rhythm notion is dependent on events separated by durations~\cite{Lacerda}. These events can be heard individually if spaced by at least $50-63ms$. To make the temporal separation between them be perceived as duration, the period should be large enough, around $100ms$~\cite{microsound}.  
It is possible to summarize the duration transitions heard as rhythm and pitch,
in the following way~\cite{Alfaix, microsound}.

\begin{table*}[htpq!]
\tiny
\centering
\caption{Transition of durations heard individually until it turns into pitch.}
\begin{tabular}{  l | r r r r   r r r    r r r || r r || r r r r r r }
\hline
           & \multicolumn{10}{c}{$\underleftarrow{\text{\bf perception zone of
           duration as rhythm}}$} & \multicolumn{2}{c}{transition} & \multicolumn{3}{c}{-} \\
duration (s) & {\bf ...}     & {\bf 32,}     & {\bf 16,}   & {\bf 8,}  & {\bf 4,}   & {\bf 2,}   & {\bf 1,}   & {\bf 1/2,} & {\bf 1/4,} & {\bf 1/8,} & $\frac{1}{16}=62,5ms$ , & $\frac{1}{20}=50ms$ & {\color{gray} 1/40} & {\color{gray} 1/80  } & {\color{gray} 1/160 } & {\color{gray} 1/320 } & {\color{gray} 1/640 } & {\color{gray} ... } \\
frequency (Hz) & {\color{gray} ...} & {\color{gray} 1/32,}   & {\color{gray} 1/16,} & {\color{gray} 1/8,} & {\color{gray} 1/4,} & {\color{gray} 1/2,} &  {\color{gray} 1,}  & {\color{gray} 2,}   & {\color{gray} 4,}   & {\color{gray} 8,}    & 16,  & 20   & {\bf 40}   & {\bf 80}   & {\bf 160}   & {\bf 320}   & {\bf 640}   & {\bf ...} \\
           & \multicolumn{10}{c}{ - } & \multicolumn{2}{c}{transition}
           & \multicolumn{6}{c}{$\overrightarrow{\text{\bf perception zone of
           duration as pitch}}$} \\
\hline
\end{tabular}
\label{tab:duracoes}
\end{table*}

The duration region marked as transition is presented minimized because the limits
are not well defined. In fact, the duration where someone begins to perceive a
fundamental frequency or a separation between occurrences, depends on the
listener and sound characteristics~\cite{microsound,Roederer}.

The rhythmic metric is commonly based on a basic duration called pulse. The
pulse typically comprehends durations between $0.25-1.5s$ ($240$
and $40 BPM$, respectively). In music education and cognitively studies, it is common to
associate these range of pulse frequencies to the durations of the heart beat,
movements of inspiration and expiration and walk steps~\cite{Lacerda,Roederer}.

The pulse is subdivided into equal parts and repeated in sequence. These
relations (division and concatenation) usually follow relations of small
order integers\footnote{In ascending order of occurrence in written and ethnic
music, these indicate the musical pulse divisions and their sequential grouping at
time: 2, 4 and 8, after that 3, 6 (two groups of 3 or 3 groups of 2) and 9 and
12 (three and 4 groups of 3). At last, the prime numbers 5 and 7, complementing
1-9 and 12. Other metrics are less common, like divisions and grouping in 13,
17, etc, and are mainly used in contexts of experimental music and classical
music of XX and XXI. No matter how complex they seem, metrics are commonly
compositions and decompositions of 1-9 equal parts~\cite{Gramani,Roederer}.}.
A schematic exposition is shown in Figure~\ref{fig:pulsoSubAgl}.

\begin{figure}[h!]
    \centering
        \includegraphics[width=\columnwidth]{figures/metricaMusical}
    \caption{Divisions and grouping in bars for the music pulse to establish a rhythmic
        metric. Divisions of the quarter note, established as the
        pulse, is presented in the left. The time signature whose specifies the same
        metrics but having as scale the groupings of the music pulse is presented in the right.}
        \label{fig:pulsoSubAgl}
\end{figure}

Dual relations (common time and binary divisions) are frequent in dance rhythms
and celebrations. Ternary relations are common in
ritualistic music and is related to the sacred. Dual relations are called imperfect, while and the ternary relations are called perfect.

The strong unities (accents) are the ones that fall in the 'head' of bars, the downbeats. The
head of a unity is the first part of the subdivision. In binary divisions (2, 4
and 8, for example), the unities originally strong turn into weak (não sei se isso ficou claro)
(e.g.\ division in 4 is: strong, weak, strong medium, weak). In ternary divisions
(3, 6 and 9) two weak unities succeeds the downbeat (e.g.\ division in 3 is:
strong, weak, weak)\footnote{Division in 6 is considered compound but can also
occur as a binary division. A binary division which suffers a ternary division
results in two units divided in tree units each: strong (subdivided in strong,
weak, weak) and weak (subdivided in strong, weak, weak). Another way to perform
the division in 6 is applying ternary division that suffers a binary division,
resulting in: a strong unit (subdivided in strong and weak) and two weak units
(subdivided in strong and weak for each).}

The accent in the weak beat is known as the backbeat, whereas a note starting in a weak
beat which duration coming across the strong beat is known as syncope.

Notes can occur inside and outside of these \emph{'music metric'} divisions. In
most behaved cases, notes occur exactly on these divisions, with broader
incidence on attacks of strong beats.
In extreme cases, the time metric cannot be perceived~\cite{Roederer}. 
Small grid variations help to build musical interpretation and differences
between styles~\cite{Cook}.

% ficou um tanto confuso:

Being pulse the grouping level $j=0$, level $j=-1$ the first pulse subdivision,
level $j=1$ the first pulse agglomeration and so on. In this way, $P_i^j$ is the
$i$-th unit of pulses at grouping level $j$: $P^0_{10}$ is the tenth pulse,
$P^{1}_3$ is the third pulse grouping unit (and, possibly, the third measure),
$P^{-1}_2$ is the second unit of pulse subdivision.

Special attention should be given to the limits of $j$: pulse divisions are
durations sensible as rhythm. Furthermore, the pulse joints sums, at its
maximum, a music or a cohesive set of musics. In other words, a duration given
by $P^{min(j)}_i$, $\forall \; i$, is greater than $50 ms$ and the durations
summed together $\sum_{\forall i}P^{\text{max}(j)}_i$ are less than a few
minutes or at most, a few hours.

Each level $j$ has some indexes $i$. A index $i$ always has three different
values (or multiple of three) that occur in a perfect relation. When $i$ has only
multiple of two, four or eight different values, than an imperfect relation occurs,
as shown in Figure~\ref{fig:pulsoSubAgl}.

Any unit (note) of a given musical sequence which has a time metric can be unequivocally
specified as:

\begin{equation}
P^{ \{ j_k \} }_{ \{ i_{k} \}}
\end{equation}

\noindent where $j_k$ is the grouping level and$i_k$ is the unit order itself.

As example, consider $P^{-1,0,1}_{3,2,2}$ as the third subdivision $P^{-1}_3$ of the
second pulse $P^0_2$ and of the second pulse group $P^1_2$. Each unit or unit set
$P_i^j$ can be associated with a sequence of temporal samples $T_i$ that forms a
music note.

The music piece \emph{Poli Hit Mia} explorers different metrics (avaleble in Appendix~\ref{ap:poli} and as part of \massa).

\subsection{Repetition and variation: motifs and large units}\label{subsec:motivos}

Given the basic music structures, either frequential (chords and scales) or
rhythmic (beat divisions and simple, compound and complex grouping), it is
natural to organize these structures in a cohesive way~\cite{Boulez}. To this end,
the concept of arcs are fundamental: we make an arc by departing from a place (music unit) and going to another one. The audition of melodic and harmonic lines is pervaded as
musical arcs thanks to the cognitive nature of the musical hearing. It is
possible to consider the note as the smaller arc, and each motif (definir motif aqui? nao foi falado ainda) and melody as another arc. Each time, beat subdivision, measure and music
section constitutes an arc. A music in which the arcs do not present consistency by
themselves, can be understood as a music with no cohesion. The coherence feeling
comes, mostly, from the skilled manipulation of arcs in a music piece.

Musical arcs are abstract structures and liable of basic operations. An spectral
arc, like a chord, can be inverted, enlarged and permuted, to cite some
examples. Temporal arcs, like a melody, a motif, a measure or a note are also
capable of variations. Remembering that
$S_j=\left\{s_j=T_i^j=\{t_i^{j}\}_0^{\Lambda_j-1}\right\}_0^{H-1}$ is a sequence
of $H$ musical events $s_j$, each event with its $\Lambda_j$ samples $t_i^j$
(refer to the beginning of this section~\ref{notasMusica}), the basic techniques
can be described as:

\begin{itemize}
        \item Temporal translation is a displacement
    $\delta$ of a specific material to another instant $\Gamma'=\Gamma + \delta$
    of the music. It is a variation that changes temporal localization in
    a music:
    $\left\{s_j'\right\}=\left\{s_j^{\Gamma'}\right\}=\left\{s_j^{\Gamma+\delta}\right\}$
    where $\Gamma$ is the duration between the beginning of the piece (or considered
    period) and the first event $s_0$ of original structure $S_j$. Observe that
    $\delta$ is the time offset of the displacement.

    \item Temporal expansion or contraction is a change in duration of each
    arc by a factor $\mu\,:\; s_j'^{\Delta}=s_j^{\mu_j . \Delta}$. Possibly,
    $\mu_j=\mu$ is constant.

    \item Temporal reversion consists of generating a sequence with elements
    in reverse order of the original sequence $S_j$, thus: $S_j'=\left\{s_j'\right\}_0^{H-1}=\left\{s_{(H-j-1)}\right\}_0^{H-1}$.

    \item Pitch translation is a displacement $\tau$ of a material for a
        different pitch $\Xi'=\Xi + \tau$. It is a variation that changes pitch
        localization of material:
        $\left\{s_j'\right\}=\left\{s_j^{\Xi'}\right\}=\left\{s_j^{\Xi+\tau}\right\}$
        where $\Xi$ is the pitch of a period $S_j$ or of an event $s_0$ of the
        original structure $S_j$. Observe that $\tau$ is the pitch offset of the
        displacement. If $\tau$ is given in semitones, the displacement in
        frequency is $\tau_f=f_0.2^{\frac{\tau}{12}}$ where $f_0$ results from
        the adoption of some reference: $f_0=\Xi_{f_0}Hz\;\sim \Xi_0$ absolute
        value of pitch. For the frequency of any pitch value: $\Xi_f=\Xi_{f_0}.2^{\frac{\Xi-\Xi_0}{12}}$. 
        For the pitch of any frequency value: $\Xi=\Xi_0 +12
        . \log_2\left(\frac{\Xi_f}{\Xi_{f_0}}\right)$.  
        In the MIDI protocol, $\Xi_{f_0}=55Hz$ while pitch $\Xi_0=33$ marks
        a \textit{A 1}, another reference point is $\Xi_{f_0}=440Hz$ which
        corresponds to $\Xi_0=69$. The unit difference is equal for semitones.
        It is not possible to say that the unit is the semitone because $\Xi=1$
        is not a semitone, but a note with an audible frequency as rhythm, with
        less than 9 occurrences each second (see Table~\ref{tab:duracoes}).

        \item Interval inversion is the inversion traversed by
        the material. The inversion is strict if the number of semitones is
        being used as a reference for the operation:
        $S_j'=\{s_j'\}_0^{H-1}=\left\{s_j^{-\varepsilon_j . f_0}\right\}$, where
        $\varepsilon_j$ is the factor between the event frequency $s_j$ and the
        frequency of $s_0$. The inversion is said tonal if the distances are
        considered in terms of degree number of the scale $E_k$:
        $S_j'=\{s_j'\}_0^{H-1}=\left\{s_j^{-\varepsilon^{\left(e_{\left(j_e\right)}\right)}
        . f_0}\right\}_0^{H-1}$ where $j_e=e_{j}^{inv}$ is the index $k=j_e$ in
        $E_k$ of the note of event $s_j$.

        \item Rotation of musical elements is the transference of the last
        element to the position of the first element, and the displacement of
        that to the penult, one position ahead. It is possible to define
        rotation of $\tilde{n}$ positions by $S'_n=S_{(n+\tilde{n})\%H}$. If
        $\tilde{n}<0$, it is enough to use $\tilde{n}'=H-\tilde{n}$. It is
        reasonable to associate $\tilde{n}>0$ with the clockwise rotation and
        $\tilde{n}<0$ with tne anti clockwise rotation. More information about rotations in presented in subsection~\ref{estCic}.

        \item The insertion and removal of materials into structure $S_j$ can be
    ornamental or structural: $S_j'=\{s_j'\}=\{s_j \text{ if condition A,
    otherwise } r_j\}$, for any music material $r_j$, including the empty
    instant. Elements of $R_j$ can be inserted at the beginning, like a prefix
    in $S_j$; at the end, as a suffix; or at the middle, dividing $S_j$ or making
    it the prefix or suffix. Both materials can be mixed in a variety of ways.

    \item Changes in articulation, orchestration and spatialization, 
    $s_j'=s_j^{*_j}$, where $*_j$ is the new characteristic incorporated by the
    element $s_j'$.

    \item Accompaniment. Both orchestration and melodic lines presented when
    $S_j$ occurs can suffer modifications and be considered as a variation of $S_j$ itself.
\end{itemize}

Other proceesings can be derived together with the presented ones, as, for example, the retrograde
inversion, temporal contraction with an external suffix, etc. The music
structures resound in the human cognitive system due to the own nature of
thought. In its many veneers, an idea read with the same number of elements and
connective aspects between them. The music, when tunned with those mental structures,
raises impressions. In this way, a whole process of mental and neurological
resonance is unleashed, responsible by the feelings, memories and imaginations,
typical to a mindful music listening. This cortical activity helps the
musical therapy, known by its utility in cases of depression and neurological
injury. It is known that regions of the human brain responsible by
hearing processing, are also used for other activities, even linguistic and mathematical~\cite{Sacks,Roederer}.

The paradigmatic structures guide the creation of new music material. One of the
most central of them is the dipole tension/relax. Traditional dipoles
are relate to tonic/dominant, repetition/variation,
consonance/dissonance, coherence/rupture, symmetry/asymmetry,
equality/difference, arrival/departure, near/far, stationary/moving,
etc. The ternary relations tend to relate to the circular and unification. The
lucid ternary communion, 'modus perfectus', opposes to the passionate
dichotomic, 'modus imperfectus'. Hereafter, there is an exposition dedicated to
directional and cyclic arcs.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Directional structures}\label{subsec:dir}

The arcs can be decomposed in two convergent sequences: the first arc reaches the
apex and the second arc returns from the apex to the start region. This apex is called climax by traditional music theory. It is
possible to differ between arcs whose climax are placed at the beginning, middle, end,
and at the first and second half of the considered duration. These structures are
shown in Figure~\ref{fig:climax}. The parameter of variation does not exist, thus
the arc consists only of the reference structure~\cite{Schoenberg}.

\begin{figure}[h!]
    \centering
        \includegraphics[width=\columnwidth]{figures/climax}
        \caption{Canonical distinctions of musical climax in a given melody and
        other domains. The considered possibilities are: climax at beginning,
        climax at first half, climax at middle, climax at second half and climax
        at the end. The y-axis is not properly specified since there exist no
        real parametric variation and, therefore, the structure is a reference.}
        \label{fig:climax}
\end{figure}

Consider $S_i=\{s_i\}_0^{H-1}$ as an increasing sequence. The sequence
$R_i=\{r_i\}_0^{2H -2}=\left\{s_{(H-1-|H-1-i|)}\right\}_0^{2H-2}$ is a sequence
that presents a perfect profiteer symmetry, i.e.\ the second half is the
mirrored version of the former. Following the musical concepts, the climax is
localized in the middle of the sequence. It is possible to modify this
fact using sequences with different sizes. All the mathematical theory of
sequences, already well established and taught routinely in calculus courses, can be
used to generate those arcs~\cite{Guidorizzo,Schoenberg}. 
Theoretically, when applied to any characteristic of music events, these sequences produce arcs, since they imply the deviation and return of an initial
parametrization. Henceforth, it is possible to have a
number of distinct arcs, with different sizes and climax for a same sequence of events. This is an interesting and useful resource, and the correlation of arcs results in coherence~\cite{Salzer}.

Historically (and practically nowadays), the golden ratio and the Fibonacci sequence have special importance. The Lucas sequence allows the generalization of
Fibonacci sequence, making it easy to understand. Given any two numbers $x_0$
and $x_1$, the Lucas sequence can be obtained as: $x_n=x_{n-1}+x_{n-2}$. The greater $n$ is, the more $\frac{x_{n}}{x_{n_1}}$ approaches the golden ratio:
$1.61803398875...$. The sequence converges fast even with high discrepant
initial values. Being $x_0=1$ and $x_1=100$, and $y_n=\frac{x_n}{x_{n+1}}$ an
auxiliary sequence, the error occurring in the first values with
relation to the golden ratio is, approximately, $\{ e_n \}
=\left\{100\frac{y_n}{1.61803398875}-100 \right\}_1^{10}=\{6080.33, -37.57, 23,
-7.14, 2.937, -1.09, 0.42, -0.1601, \\ 0.06125, -0.02338\}$. The Fibonacci sequence
presents about the same error progression but starts right at the second step since $\frac{1}{1}\approx\frac{101}{100}$.

The music piece \emph{Dirracional} exposes the use of arcs into directional structures.
Its source code is included in Appendix~\ref{ap:dirracional} and available online as part
of \massa~\cite{MASSA}.

\subsection{Cyclic structures}\label{estCic}

The philosophical understanding that the human thought is based on the perception of
similarities, differences, stimuli and objects, places the symmetries
at the core of the cognitive process~\cite{Deleuze}. Mathematically, symmetries
are algebraic finite groups, and a finite group is always isomorphic to a permutation
group. It is possible to say that permutations represent any symmetry in a
finite system (precisa de uma referência aqui). In music, permutations are ubiquitous and considerably present in techniques, which confirms their central role. The successive application of permutations generates cyclic arcs~\cite{change,Zamacois,permMusic}. Two academic studies were dedicated to this end, aiming at generating music structures~\cite{figgusOriginal, figgusEspacializacao}.

Any permutation set can be used as a generator of algebraic groups~\cite{permMusic}.
The properties defining a group $G$ are:

\begin{multline}\label{eq:groups}
%\begin{split}
\forall \;\; p_1,p_2 \in G \Rightarrow\quad\;\; p_1 \bullet p_2  =
p_3 \in G \\ \;\;\;\text{(closure property)} \\
\forall \;\; p_1,p_2,p_3 \in G \Rightarrow\quad (p_1\bullet p_2)\bullet p_3  =
p_1\bullet (p_2\bullet p_3)\quad\;  \\ \text{(associativity property)} \\
\exists \;\; e \in G :\; p \bullet e  = e \bullet
p \;\;\;\; \forall p \in G  \quad \\ \text{(identity element property)} \\
\forall \;\; p \in G, \;\exists\; p^{-1} :\quad\;  p\bullet p^{-1} 
=p^{-1}\bullet p = e \\ \; \text{(inverse element property)}
%\end{split}
\end{multline}

From the first property it is possible to conclude that any permutation can be
operated with another permutation. In fact, it is possible to apply a
permutation $p_1$ and another permutation $p_2$, and, comparing both initial and final
orderings, results in another permutation $p_3$.

Every element $p$ operated with itself a sufficient number of times $n$ reaches
the identity element $p^n=e$ (otherwise the group would be infinite, generated
by $p$). The lower $n\;:\;p^n=e$ is called the element order. Thus, a finite
permutation $p$, successively applied, reaches the initial ordering of its
elements, making a cycle. This cycle, if used as parameter of music notes,
implies in a cyclic musical arc.

These arcs can be established by using a set of different permutations. As a historic
example, the tradition called \emph{change ringing} conceives the music through
bells played one after other and then played again, but in a different
order. This process repeats until it reaches the initial ordering. The set of
different traversed orderings is a \emph{peal}. Table~\ref{tab:change}
represents one traditional \emph{peal} for 3 bells (1, 2 and 3) which explores
all possible orderings. Each line indicates one bell ordering to be
played. Permutations occur between each line. In this case, the music structure
consists of permutations by itself and some different permutations operate due to
the cyclic behavior.

\begin{table}[htpq!]
\centering
\caption{Change Ringing: Standard \emph{peal} for 3 bells. Permutations
occur between each line. Each line is a bell ordering and each ordering is played
        a line a time.} 
\begin{tabular}{l c r}
\textcolor{red}{1} & \textcolor{blue}{2} & \textcolor{green}{3} \\
\textcolor{blue}{2} & \textcolor{red}{1} & \textcolor{green}{3} \\
\textcolor{blue}{2} & \textcolor{green}{3} & \textcolor{red}{1} \\
\textcolor{green}{3} & \textcolor{blue}{2} & \textcolor{red}{1} \\
\textcolor{green}{3} & \textcolor{red}{1} & \textcolor{blue}{2} \\
\textcolor{red}{1} & \textcolor{green}{3} & \textcolor{blue}{2} \\
\textcolor{red}{1} & \textcolor{blue}{2} & \textcolor{green}{3}
\end{tabular}
\label{tab:change}
\end{table}

The use of permutations in music can be summarized in the following way:
consider $S_i=\{s_i\}$ as a sequence of musical events $s_i$ (e.g.\ notes), and a
permutation $p$. $S_i'=p(S_i)$ comprises the same elements of $S_i$ but in a
different order. The permutations can be written based on two notations: cyclic or
natural. The natural notation basically comprehends the order of the resulting indexes from
the permutation. Thus, agreed the original ordering given the sequence of its
indexes $[0\;1\;2\;3\;4\;5\;...]$, the permutation is noted by the sequence it
produces (ex. $[1\;3\;7\;0\;...]$). In the cyclic notation, permutation is the
changing of one element by the posterior one, and the last element by the first one.

It is not necessary to permute elements of $S_i$, but only some
characteristics. In this way, being $p^f$ a permutation in frequency and $S_i$ a
sequence of basic notes as exposed in the end of section~\ref{notaBasica}, the
new sequence $S_i'=p^f(S_i)=\left\{s_i^{p(f)}\right\}$ consists of the same
music notes, respecting same order and maintaining the same characteristics, but with the
fundamental frequencies permuted following the pattern that $p^f$ presents.

It is worthwhile to mention two subtleties of this procedure. First, the permutation $p$ does not have to involve all elements of $S_i$, i.e.\ it can operate in subsets of $S_i$. Second,  the elements $s_i$ do not have to be entirely executed at each state access. To exemplify,
consider $S_i$ as a sequence of music notes $s_i$. If $i$ goes from $0$ to $n$, and
$n>4$, at each measure of $4$ notes it is possible to execute the first $4$
notes. The other notes of $S_i$ can occur in measures where permutations set
aside those notes to the first four notes of $S_i$.

Each one of the permutations $p_i$ described above relates the note dimensions where it operates (frequency, duration, \emph{fades},
intensity, etc) and the period of incidence (at how much access a permutation is
applied). During realization of notes in $S_i$, an easy and coherent form is to
execute the first $n$ notes\footnote{The execution of disjoint notes of $S_i$
equals to modify the permutation and to execute the first notes.}.

Appendix~\ref{cap:FIGGUScode} and \massa present the computational implementation about permutations~\cite{MASSA,figgusOriginal,figgusEspacializacao}.

\subsection{Musical idiom?}

There are many studies intending to model and explore a 'musical language',
or 'linguistics applied to music' or even to discern between different 'musical
idioms'~\cite{Lerdahl, Harmonia, Salzer,Alfaix}. In a simple way, a musical idiom
is the result of a chosen material together with repetition of elements and
repetition of relations between elements present along the music. The related dichotomies are outstanding, as explained at subsection~\ref{subsec:motivos}: repetition and variation, relax and tension, equilibrium and instability, consonance and dissonance, etc.

\subsection{Musical usages}\label{subsec:usosmusicais3}

The basic note was defined and characterized in quantitative terms (section~\ref{sec:notaDisc}). Next, the internal note
composition was addressed and both internal transitions and immediate treatment
were understood (section~\ref{varInternas}). Finally, this section aims at
organizing these notes in music. The gamma of resources and consequent infinitude
of possibilities is a typical situation and highly relevant for art contexts~\cite{Harmonia,Webern}.

There are studies dedicated to each one of the presented resource. For example, it is possible to obtain the 'dirty' triadic harmonies (with notes without the triad) by
superpositions of dirty fourths. Another interesting example is the simultaneous
presence of rhythms in different metrics, constituting what is
called \emph{polyrhythms}. The music piece \emph{Poli-hit mia} explores these
simultaneous metrics by impulse train (??) convolved with notes that compound the
line. Its source code is included in Appendix~\ref{ap:poli} and available online as part
of \massa.

The microtonal scales are important for the music of the 20th
century~\cite{microtonalidade} and present diverse outstanding results, like the
fourths of tone in the Indian music. The musical sequence \emph{MicroTom}
explores these resources, including microtonal melodies and microtonal harmonies
with many notes in a very reduced note scope. Its code is presented in
Appendix~\ref{ap:micro} and available online in \massa.

As pointed in subsection~\ref{subsec:mus2}, the relations between
parameters are powerful ways to acquire music pieces. The number of permuted
notes can vary during the music, revealing relationship with piece
duration. Harmonies can be made of triads (eqs.~\ref{triades}) with replicated
notes at each octaves and more numerous as minor the depth and frequency of
vibratos (eqs.~\ref{vbrGamma},~\ref{vbrAux},~\ref{vbrF},~\ref{vbrGamma2},~\ref{vbrT}),
among other uncountable possibilities.

The presented symmetries at octave divisions (eqs.~\ref{escSim}) and the
symmetries presented as permutations (Table~\ref{tab:change} and
eqs.~\ref{eq:groups}) can be used together. In the music piece \emph{3 trios}
this association is done in a systematic way in order to enable the required
listening. This is a instrumental piece, not included as a source
code~\cite{3Trios}.

The \emph{PPEPPS} (Pure Python EP: Projeto Solvente) is an EP synthesized using
the resources presented in this study. With minimal parametrization, the program
generates complete musics, allowing the easy composition of musics and sets of
musics. Using few lines of code it is possible to obtain
a directory with the generated musics. This facility and technological
demystification create aesthetically possibilities for both sharing and education.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions and future works} %Nome do capítulo.
\label{cap:conclusao}

We have described a concise system that explores the musical elements and relate them to the digital signal. The reader is invited to access \emph{Scripts} and \massa, where these relations are computationally implemented. We hope the didactic report along the paper and the supplied scripts facilitates and encourages the use of the proposed framework.

The open possibilities provided by the techniques and results discussed in this paper involve psychoacoustic experiments and the creation of interfaces for the generation of music, noise and other music applications in a high fidelity (\emph{hi-fi}). It is worthwhile to mention the benefit of these results for artistic and didactic purposes. The incorporation of programming skills is facilitated by the provided visual aids. Initial \emph{live-coding} practices and guided courses based on this work have already been realized with successful. Examples are Puredata and Chuck.

This work systematically investigates the parameterization issues (like the tremolo, ADSR, etc.) in a high fidelity, which has significant artistic utility. Such detailed analytical descriptions, together with the computational implementations, have not been covered before in the literature (as shown in Appendix~\ref{cap: trabalhosRelacionados}).

Besides, the free software license and online availability of the exposed content as hypertext, in conjunction with the respective codes and sound examples, strongly facilitates future collaborations and the generation of sub-products in a co-authorship fashion. As consequence, the expansion of \massa is feasible and straightforward, making it possible new developments and implementations of musical context.

In addition, this work permitted the formation of groups with common interests, like music creativity and computer music. Specially, the project \url{labMacambira.sf.net} groups Brazilian and foreign co-workers in order to offer relevant contributions in diverse areas like Digital Direct Democracy (não devo ter traduzido certo), georeferencing techniques, artistic and educational activities. Some of these reports are available in Appendix~\ref{cap:musicaExtra} and made online. There are more than 700 videos, written documentations, original softwares and contributions in well-known external softwares such as Firefox, Scilab, LibreOffice, GEM/Puredata, to name a few~\cite{siteLM,wikiLM,vimeoLM}.

Future works include the application of the reported results in machine learning and artificial intelligence methods for the generation of appealing artistic materials.

%% Há um aumento no número de pesquisas relacionadas à música em andamento no campus de São Carlos da USP, o que sugere facilidade para estabelecer parcerias.
%% As publicações acadêmicas efetivadas durante este mestrado também apontam para uma multidisciplinaridade,
%% tratando diretamente de questões humanas como artes, filosofia, humor e linguagem falada e escrita, através de artifícios lineares e estatísticos.\cite{FabbriSTAT,FabbriACL,FabbriComplenetVoz,FabbriComplenetTexto} Os desdobramentos estão alcançando redes sociais e teorias epistemológicas com base em pesquisas prévias, dos orientadores deste trabalho, com forte presença de redes complexas e processamento de linguagem natural.

\nocite{*}
\bibliography{phi+mus}
\end{document}