### 2017-07-30 article/acm/listings.tex
iniciado documento para as listagens dos elementos no artigo

\documentclass{scrreprt}
% \documentclass[12pt]{article}
\usepackage{graphicx}

\usepackage{xr}
\externaldocument{sample-acmsmall}

\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{hyperref}
\hypersetup{
        colorlinks,
        linkcolor={red!50!black},
        citecolor={blue!50!black},
        urlcolor={blue!80!black}
}
%
\usepackage{graphicx} %allows for including images
\usepackage{biblatex}

\newcommand{\massa}{{\large \textsc{mass}}}

\makeatletter
\newcommand{\usetocfromothersource}[1]{%
  \begingroup
  \IfFileExists{#1}{%
    \@input{#1}% 
    \@nobreakfalse
  }{}%
  \endgroup
}
\makeatother

\def\thesection{SI-\arabic{section}}

\begin{document}

\title{Listings of equations, figures, tables and sections of the
article '{\it \bf Musical elements in the discrete-time representation of
sound}' and of the scripts in the MASS toolbox}

\author{\\\\\\ Renato Fabbri (USP),\\
Vilson Vieira da Silva Junior (The Grid),\\
Ant\^onio Carlos Silvano Pessotti (Unimep),\\
D\'ebora Cristina Corr\^ea,\\
Osvaldo N. Oliveira Jr. (USP)
}

\maketitle

\begin{abstract}
The article is the main document of the \massa\ toolbox.
Being it of considerable length and complexity,
this document contains listings of its elements to facilitate
its navigation, apprehension and general usage.
\end{abstract}

\tableofcontents

\clearpage
\section{Table of Contents of the article}
\makeatletter
\let\Hy@linktoc\Hy@linktoc@none
\makeatother
\usetocfromothersource{sample-acmsmall.toc}

\clearpage
\section{Equations}

\clearpage
\section{Figures}

\clearpage
\section{Tables}

\clearpage
\section{Scripts}
\subsection{For all equations in each chapter}
\clearpage
\subsection{To render musical pieces}
\clearpage
\subsection{To render the figures used in the article}

\clearpage
\section{Other documents}
Latex files. PDF files.
MSc Dissertation.

\section{Final considerations}

\end{document}